### 2012-12-04 paper.tex
First pass on the paper structure. Only empty sections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

%\documentclass{article}
\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\usepackage[utf8x]{inputenc}
\usepackage{cite}
\usepackage{graphicx}
\graphicspath{{figs/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

\usepackage[draft]{fixme}

\title{\LARGE \bf
Explicit Knowledge and the Deliberative Layer: Lessons Learned
}

\author{Séverin Lemaignan and Rachid Alami\\
CNRS-LAAS, 7 av. du Colonel Roche, F-31077 Toulouse, France\\
Université de Toulouse, UPS, INSA, INP, ISAE, LAAS, F-31077 Toulouse, France\\
{\tt surname.name@laas.fr}
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Knowledge-Oriented Architecture}

\subsection{The LAAS deliberative layer}

\cite{Alami2011}

\begin{figure*}
        \centering
        \includegraphics[width=1.7\columnwidth]{archi}
        \caption{Overview of the LAAS deliberative layer.}
        \label{fig|archi}
\end{figure*}

\subsection{Knowledge Model}

\cite{Lemaignan2010}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Situation Assessment}
\label{sect|sit-ass}

\begin{figure}
        \centering
        \includegraphics[width=\columnwidth]{spark-oro}
        \caption{Functional overview of knowledge base (\emph{oro-server}, top part) and the geometric situation assessment module (\emph{SPARK}, bottom part)}
        \label{fig|spark-oro}
\end{figure}

\cite{Sisbot2011}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Communication}
\label{sect|com}

\cite{Lemaignan2011a}

\subsection{Disambiguation at semantic level}

\cite{Ros2010b}

\fxnote{Mention failures like 'placeOing'}

\subsection{Multi-modal communication}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robot Control}
\label{sect|ctrl}

\subsection{Modeling the interaction}

We split the interaction situations stemming from the situation assessment and
communication components in two categories: \emph{desires} and
\emph{experiences}.

\subsection{Task planning}

\cite{Alili2008}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Internal Cognitive Processes}
\label{sect|intern}

\subsection{Theory of Mind}

\cite{Warnier2012a}

\subsection{Working Memory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sect|conclusion}

\subsection{The Architect view: loose coupling and modalities merging}

\subsection{The Logician view: the importance of trivial inferences}

\subsection{The Cognician view: palpable knowledge and semantic thinking}

\section*{Acknowledgment}

This work has been supported by EU FP7 ``SAPHARI'' under grant agreement no. ICT-287513.

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,biblio}

\end{document}

### 2012-12-04 paper.tex
Imported sections from the thesis

@@
 
+\newcommand{\ie}{{\textit{i.e.\ }}}
+\newcommand{\cf}{{\textit{cf\ }}}
+\newcommand{\eg}{{\textit{e.g.\ }}}
+
 \title{\LARGE \bf
@@
 
+Over the last four years, we have been slowly ramping up explicit knowledge
+representation and manipulation in the deliberative and executive layers of our
+robots. Ranging from situation assessment to symbolic task planning, from
+verbal interaction to event-driven execution control, we have built up a
+\emph{knowledge-oriented} architecture which is now used on a daily basis on our
+robots.
+
+This article presents our design choices, the articulations between the diverse
+deliberative components of the robot, and the strenghts and weaknesses of this
+approach. We show that explicit knowledge management is not only a convenient
+tool from the software engineering point of view, but also pushes for a
+different, more \emph{semantic} way to address the decision-making issue in
+autonomous robots.
+
 \end{abstract}
@@
 \subsection{The LAAS deliberative layer}
+
+We can give a broader look at the knowledge and the streams of knowledge in our
+systems.  Based on the experience gained while developing and deploying {\sc
+ORO}, our ontology-based knowledge server, we have presented how symbolic
+knowledge could be produced from perception and geometric reasoning in modules
+like {\sc SPARK}, a grounded, perspective-aware, geometric reasoner. We have
+seen how symbolic knowledge could be reused by different control systems and
+task planners like {\sc Cram}, {\sc SHARY}, {\sc pyRobots}, the {\sc CLSU
+Toolkit} or {\sc HATP} and how they take advantage of semantic abstractions
+provided by knowledge base. We have also presented the bidirectional
+integration of {\sc Dialogs}, a natural language processor for English, with
+the knowledge base.
+
+Altogether, these components compose an architecture that we call
+\emph{knowledge-oriented}:
+
+\begin{itemize}
+
+    \item{Knowledge is explicitly stored in one central and consistent
+    repository of facts, accessible by all modules.}
+
+    \item{Knowledge is represented in a strict formalism (OWL statements) and
+    with a clearly defined vocabulary (stated in the common-sense ontology).}
+
+    \item{The first two points enable both a loosely-coupled architecture where
+    modules can very easily be removed or replaced by other ones as long as
+    they share the same semantics (modules are defined by the knowledge they
+    produce),}
+
+    \item{and a \emph{symbolic} reactive, event-driven approach to supervision.
+    By managing events at the same level as the reasoner, we take full
+    advantage of the inference abilities of ORO to trigger events whose
+    \texttt{true} conditions can be inferred.}
+
+    \item{Finally, this architecture allows for the combination of very
+    different knowledge modalities in a single homogeneous environment,
+    bringing mutual benefits to components. For instance, the dialogue
+    processing module can perfectly run without any geometric
+    perception, but its disambiguation routines can transparently
+    benefit from it when available (since richer symbolic descriptions of
+    objects are then available).}
+
+\end{itemize}
+
+This architecture moves away from standard layered approaches. Interactions
+between components are mostly bidirectional and, from the software components
+point of view, we do not introduce layers of abstraction (we do, however, have
+access to the lower level modules of the robot to execute actions, but all
+cognition-related modules reside at the same level). This is especially visible
+for the dialogue input processing. This component does not simply act as an
+alternative perceptual input to the symbolic database, but also actively
+queries previously acquired knowledge to disambiguate and validate the newly
+created symbolic knowledge.
+
+Our architecture relates but is to be distinguished from \emph{Beliefs,
+Desires, Intentions} (BDI) architectures. BDI architectures are primarily
+focused on \emph{practical reasoning}, \ie the process of deciding, step by
+step, which action to perform to reach a goal (as summarised by
+Woolridge~\cite{Woolridge1999}). The management of the interaction between
+knowledge (the beliefs) and task and plan representation and execution (the
+desires and the intentions) is central, and aims at selecting at each step the
+best subgoal. It becomes then an intention that the robot commits to.
+
+This interaction between knowledge and actions is also central to our approach
+(as for any cognitive system), but task representation and task execution is
+not seen as a monolithic, central function: it is one of the activities of the
+robot, actually split between communication components (that can acquire
+desires from interaction with agents, amongst other things) and an execution
+controller that may decide to take an incoming desire into account to create
+its own internal goals. The controller generates and controls intentions from
+these goals with the help of a symbolic task planner, that has also direct
+access to the knowledge base.
+
+The architecture is not only focused on this workload, and other
+activities are conducted without being explicitly considered as desires:
+assessment of the situation and the environment, dialogue (including
+performative dialogue that can possibly change the internal state of the robot,
+but does not lead to the creation of desires,  like question answering or
+statement assertion), various background monitoring and recognition tasks, etc.
+
+Regarding the anchoring question, this architecture is bidirectional. The
+components we described provide a \textit{bottom-up} grounding process: SPARK
+and \textsc{Dialogs} constantly build and push new symbolic contents about the
+world to ORO where it becomes accessible to decisional layers. In parallel, ORO
+relies on reasoning in a \textit{top-down} way to produce new facts that may
+trigger in return physical behaviours.
+
+We believe that this \emph{knowledge-oriented} approach has a strong potential
+not only to enable rich human-robot interaction, but also as a broader approach
+to information alignment and fusion in complex robotic systems.  The
+versatility of this paradigm could be illustrated by a simple imaginary
+scenario with a blind robot and a deaf robot. The blind robot does not see (no
+cameras or alike), but someone can verbally describe a scene to it. On the
+other hand, the deaf robot has a good vision system, but cannot process verbal
+input.  Without any changes to the software architecture that we described,
+control modules of both robots could equally perform the same tasks since all
+the knowledge is abstracted and centralised (note that to actually implement
+this imaginary situation, the blind robot would of course need \textit{a
+priori} 3D models of objects talked about to enable planning or pick and place
+actions, and the deaf robot would require at least some gesture interpretation
+to understand orders).
+
+This architecture may also contribute to bring closer robotics and psychology:
+it provides clear entry points to implement some classical psychology tests to
+robots. For instance, we presented experiments focused on issues related to
+perspective taking. By explicitly enabling independent modeling of the beliefs
+of each agent, our architecture is especially well suited to set up cognitive
+and psychological experiments that involve a theory of mind, such as
+\emph{False-Belief} experiments, as recently presented in~\cite{Warnier2012a}.
 
@@
 
+Experiences comprise of emotions and questions
 \subsection{Task planning}
@@
 
+\begin{quote}
+
+    Where to find milk? Milk is a subclass of dairy which is itself a subclass
+    of a perishable goods. The usual storage place for perishable goods is the
+    fridge, so the milk is likely to be found in a fridge.
+
+\end{quote}
+
+This example of reasoning, quoted from Moritz Tenorth, is a good example of
+simple yet non-trivial reasoning. As a matter of fact, only very few of such
+reasoning cases where positively identified in our scenarii and experiments
+(and consequently implemented as rules in ORO).
+
+The design choices of our architecture partially explain that fact: first, the
+planning task (which is the prototypical reasoning task) is delegated to a
+dedicated, external planner. Then, time is not represented in ORO, and
+consequently no temporal reasoning takes place at this level: action
+recognition or monitoring are handled by other layers, and the underlying
+reasoning tasks are not implemented as explicit symbolic rules in the knowledge
+base.
+
+The experiments we have conducted are also likely to have too simplistic
+semantics to let complex reasoning needs to emerge. Scenarii with more complex
+semantics would be desirable to better stress the expressiveness and inference
+abilities provided by description logics.
+
+Is reasoning at the knowledge level immature or even superfluous, then? Not so:
+hundreds of trivial (from a human point of view) inferences are continuously
+produced by the system (translating inheritance relations, domain/range
+constraints, transitivity, etc.) and encode a large amount of common-sense
+knowledge that would be tedious, to say the least, to manually assert. These
+trivial inferences are all the more important that an expressive knowledge
+representation language is used: when a language like OWL allows to directly
+represent high-level concepts like partitions, cardinality restrictions,
+properties' ranges and domains, it leads to a more implicit (because more
+abstract) description of the vocabulary that in turn requires more underlying
+reasoning. With the progress in the understanding of the relations between
+expressiveness and (tractable) satisfiability, along with the progress of
+reasoners, more and more of the inferences do not need to be explicit anymore,
+and consequently move behind the scenes.
+
+And we predict that \emph{common-sense encoding} is likely to remain the main
+application of reasoning in our robotic architectures, where reasoning related
+to decision making mostly happen outside the knowledge representation system.
+
 \subsection{The Cognician view: palpable knowledge and semantic thinking}
+
+The main motive goal was to transform the knowledge in the robot from some ubiquitous,
+pervasive, multi-modal and, most importantly, mostly undefined feature of the
+system into an observable, quantifiable, manipulable resource, what we could
+call a \emph{palpable} feature.
+
+This transformation, both from the technical point of view (the ORO server, the
+ontologies, the bindings, etc.), and as a more subtle change in the practises
+related to the development of robotic components, is the main contribution of
+this thesis.
+
+Knowledge is not an abstract concept anymore: it is a set of statements, in
+most cases directly intelligible to the developers, stored in one place. We can
+export them, monitor them, review them, question them.
+
+Communication between the robot's module is now conceived in term of what are
+the \emph{semantics} of the information flows, instead of a simple
+compatibility of interfaces. When defining the frontiers of a robotic
+component, we do not think anymore only in term of ``is the interface complete
+and self-contained'', but also in term of ``is the semantic complete and
+consistent?''. This allow a deeper, more correct modularity: two modules that
+share the same, well-defined semantic can be confidently exchanged. When we
+remove or disable a component (the dialogue processing, the geometric
+reasoning, ...), we know precisely what knowledge will not be available
+anymore.
+
+We call this new property of our robot, that allows for both qualitative and
+quantitative analysis of the beliefs, its \emph{cognitive observability}.
+
+It is somewhat related to the idea of \emph{cognitive penetrability} introduced
+by Pylyshyn~\cite{Pylyshyn1989} in 1989, in the context of the study of
+possible strong equivalences between computational models and the
+\emph{psychological reality}:
+
+\begin{quote}
+
+    [One of the criterion] relies on the assumption that we can identify
+    certain clear cases of phenomenon that should be accounted for at the
+    knowledge level, that is, in terms of the representations alone, rather
+    than in terms of properties of the cognitive architecture. Phenomena that
+    depend in a rational way on subjects' goals, beliefs, and utilities are a
+    case in point. For example in psychophysics we assume that if a measure
+    (such as a threshold) changes systematically as we change the payoffs (that
+    is, the relative cost of errors of commission and of omission), then the
+    explanation of that change must be given at the knowledge level -- in terms
+    of decision theory -- rather than in terms of properties of sensors or
+    other mechanisms that are part of the architecture. In general showing that
+    certain empirical phenomena are sensitive to goals and beliefs (or what I
+    call \emph{cognitively penetrable}) is prima facie evidence that they
+    should not be attributed to properties of the architecture.
+
+\end{quote}
+
+The introduction of an explicit \emph{knowledge level} in our architecture
+makes it possible to effectively assess the cognitive penetrability of the
+whole robot behaviours (this is however not new, and traditional BDI
+architectures would also make this claim).
 

### 2013-03-04 paper.tex
Written part 'Architect view' of the conclusion

@@
 
+We conclude this article we three different perspective on the knowledge-centric
+techniques we have presented here: first, the \emph{architect perspective}:
+what are the pro and cons of our approach from an engineer point of view. Then,
+the \emph{logician perspective}: does the promise of automated reasoning for
+robots really comes to fruition with ontologies? And eventually the
+\emph{cognitician perspective} or how explicit knowledge manipulation endows
+our robots with new cognitive capabilities.
+
 \subsection{The Architect view: loose coupling and modalities merging}
+
+In this article, we have presented several sub-systems of our robots: a module for geometric
+reasoning and situation assessment (written in C++), a natural
+language interface (written in Python), a symbolic task planner (written in
+C++), two execution controller (one written in Python, one in PRS). These
+modules communicate together through RDF statements with only a few exceptions.
+
+Since these statements are anchored in the robot's pool of knowledge, they
+actually convey \emph{meaning} through the system (for instance, when the
+dialog module or the task planner or the execution controller uses the symbol
+{\tt Give} \emph{means}, it refers to the abstract class of all give actions, which are also stated
+to be a kind of purposeful actions). This has one important consequence from
+the perspective of the system architecture: modules interfaces are only defined
+in term of knowledge consumption and/or production. This brings good decoupling
+properties. For instance, the output of the dialog module is a new
+\emph{situation desired} by the speaker. This module can be transparently
+replaced or completed by any other module that produces the same chuncks of
+knowledge. Only a thin protocol layer remains to get the modules to interact
+together. This allows us to implement (otherwise difficult) merging of
+radically different perception modalities like vision (the robot sees an object on
+a table) and verbal description (a human says to the robot another object is
+under the table): both are converted to RDF statements ({\tt obj1 type Object},
+{\tt obj1 isOn table1} and {\tt obj2 type Object}, {\tt obj2 isUnder table1})
+and made available to the other robot modules.
+
+Another consequence is that the communication channels are {\it
+defacto} defined by the semantics of the information they
+convey. This allows the system designers to think in term of which module
+produces or needs which knowledge, instead of which module produces which
+service. We found this approach to be very helpful when implementing cognitive
+functions like the theory of mind, the perspective taking or the natural
+language grounding.
+
+Lastly, another good property of knowledge-based communication between module
+is the \emph{cognitive observability} of the system: by logging queries between
+modules (which is even easier in our case since we rely on a central knowledge
+management), one can trace the interactions between the robot systems at a high
+abstraction level. We will come back below on the significance of this property
+from the cognitive point of view.
+
+The knowledge-centric architecture also has downsides. We can name here a few.
+
+Because the sources of knowledge the robot has to deal with are multiple and
+often not known at startup (typically, the knowledge generated by a verbal
+interaction with a human), it is difficult to formally guarantee reliability.
+
+Also, it must be noted that actuation itself is directly piloted by the
+execution controller through standard RPC calls (ROS actions and/or Pocolibs
+requests): no knowledge abstraction takes place.
+
+And obviously, many aspects like uncertainty management, modal logics,
+non-monotonic reasoning,... remain to be explored to gain the level of
+expressiveness required to cover many of the more complex human-robot
+interaction scenarii.
 

### 2013-03-07 paper.tex
Worked on the intro

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
+\usepackage{url}
 \graphicspath{{figs/}}
@@
 
-\subsection{The LAAS deliberative layer}
+\subsection{Towards the cognitive robot at LAAS}
+
+Natural interaction and cooperation are the current (dare we say,
+\emph{short-term}) targets for the human-robot interaction community. They are
+keys to the broad class of \emph{interactive manipulation problems}: several
+agents agree on a (more or less implicit) joint goal that requires some sort of
+cooperation to be successfully achieved. This class of problems involves both
+dialogue and manipulation and they are often not completely defined at
+start-up: they require iterative and interactive resolution.
+
+Over the last years we have focused our efforts on identifying the cognitive
+prerequisites of these challenges, and giving them experimental reality on the
+robots: what is required for a sentence like ``Let's dress the table together''
+to be understood by the robot, correctly interpreted in the spatial and
+temporal context of the interaction, and eventually transformed into a set of
+actions?
+
+We have chosen to tackle the challenge from several ends: human-aware
+navigation and motion planning (\cite{...Mamoun}), situation assessment coupled
+with motion planning (\cite{...}), projection of ``mightabilities'' that
+anticipates what surrounding agents may do (\cite{Amit...}), and the design and
+deployment of a pervasive \emph{knowledge-centric} software architecture.
+
+We focus in this paper on that last item, to present what that means, and what
+are the benefits and drawbacks.
+
+\subsection{Deployment on robots}
+
+While we do not present them here, the ideas and techniques pertaining to this
+paper have been all implemented and tested on several robots (see previously
+reported experiments in~\cite{Lemaignan2010, Ros2010b, Lemaignan2011a,
+Warnier2012a}). Videos presenting some of these experimental results can also
+be watched online at~\url{http://www.laas.fr/~slemaign}.
+
+\subsection{Article overview}
+
+We first describe with more details how the deliberative layer of our robots is
+built and why this monolithic functional group enables what we call a
+\emph{knowledge-oriented} architecture. The first section also introduce our
+central knowledge management tool, the ORO server.
+
+The next sections go over each of the facets of this deliberative layer: how
+knowledge is produced asynchronously from geometric reasoning, how the robot
+conducts grounded multi-modal interaction with humans, how knowledge is used by
+decision-making components like the robot controller or the task planner, and
+finally, how our knowledge-centric architecture enables the implementation of
+specific internal cognitive processes.
+
+We conclude the article with a discussion about the strenghts and weaknesses of
+such a knowledge-oriented architecture, fron three perspectives: the architect,
+the IA expert and the cognitician.
+
+\section{The LAAS deliberative layer}
+
+\cite{Alami2011}
+
+\begin{figure*}
+        \centering
+        \includegraphics[width=1.7\columnwidth]{archi}
+        \caption{Overview of the LAAS deliberative layer.}
+        \label{fig|archi}
+\end{figure*}
 
@@
 \emph{False-Belief} experiments, as recently presented in~\cite{Warnier2012a}.
-
-\cite{Alami2011}
-
-\begin{figure*}
-        \centering
-        \includegraphics[width=1.7\columnwidth]{archi}
-        \caption{Overview of the LAAS deliberative layer.}
-        \label{fig|archi}
-\end{figure*}
 

### 2013-03-07 paper.tex
Section on task planning

@@
 
-\subsection{Modeling the interaction}
+\subsection{Modeling the knowledge related to the interaction}
 
@@
 Experiences comprise of emotions and questions
+
 \subsection{Task planning}
 
-\cite{Alili2008}
+In order to devise how a given goal can be accomplished, the robot has to
+elaborate a plan,\ie a set of actions to be achieved by the robot and its human
+partners.  We use in our architecture the HATP planner~\cite{Alili2009} (for
+Human Aware Task Planner).  HATP is based on a Hierarchical Task Network (HTN)
+refinement, which performs an iterative task decomposition into sub-tasks until
+reaching atomic actions.  The planning domain defines a set of methods
+describing how to decompose a task and can be seen as the procedural knowledge
+of the robot.
+
+In order to produce a collaborative behaviour, the robot plans not only for
+itself but also for the other agents. The planning domain of each agent is
+instantiated from the agent-specific model in the ORO server. The resulting
+plan, called ``shared plan'' is a set of actions that form a stream for each
+agent involved in the goal achievement. HATP can also be tuned by setting up
+different costs depending on the actions to apply and by taking into account a
+set of constraints called \emph{social rules}. This tuning aims at adapting the
+robot's behaviour according to the desired level of cooperation of the robot.
+
+Two important remarks: because HATP is a generic symbolic task planner, we have
+been able to design a planning domain at a semantic level which is close to the
+one used in the human-robot dialogue (the planner vocabulary contains concepts
+like \texttt{give}, \texttt{table}, \texttt{is on}...). Hence only a few
+ontology rules have been required to map both the knowledge extracted from the
+situation assessment and the statements originated from the verbal interaction
+to the planner domain.
+
+Second remark, after some research (see appendix B of~\cite{ma these} for a
+detailled discussion)  we have decided not to represent neither tasks nor plans
+in the ontology: the planning domain (with tasks pre- and postconditions) is
+stored in a specific format, outside of the knowledge repository, and the
+resulting plans are directly communicated to the robot controller. The ontology
+formalism that we have adopted (namely OWL) does not appear to be a satisfaying
+reprepresentation language for concepts that are both tightly bound to time
+representation and non-monotonic. Other researches (\cite{tenorth...}) have attempted with more success (\fixme{regarder ca!!}).
 
@@
 
-We conclude this article we three different perspective on the knowledge-centric
+We conclude this article with three different perspectives on the knowledge-centric
 techniques we have presented here: first, the \emph{architect perspective}:

### 2013-03-10 paper.tex
Section on semantic disambiguation

@@
 
+\subsection{Natural language grounding}
+
 \cite{Lemaignan2011a}
@@
 
-\cite{Ros2010b}
+One protoypical example of semantic disambiguation has been given in
+\cite{Ros2010b} with the child's \emph{spygame}: two players are facing
+eachother with a set of random objects inbetween, one player mentally choose
+one object, and the other player has to guess the object by asking closed
+questions like \emph{Is your object small or large?}. Based on the knowledge it
+has acquired, the robot is able to minimize the number of questions required to
+find the object.
+
+When playing this kind of game, however, the issue arises that the robot has no
+way to select which knowledge about the object is relevant in the interaction
+context. For instance, the knowledge base may store facts like \stmt{obj1 type
+ActiveConcept} (which internally means that this concept was mentioned in a
+discussion in the last few seconds): this information is not a relevant
+property of \concept{obj1} when trying to disambiguate concepts with humans.
+This distinction between \emph{internal knowledge} (meaningful to
+the systme only) and \emph{common knowledge} (whose meaning is understood by
+all the interactors) has not been properly dealt with in our architecture.
+
+Besides, even knowledge that belongs to what we call \emph{common knowledge},
+may not be appropriate in a given interaction context. For instance, the system
+may compute that at a given instant, the human is looking at the object:
+\stmt{human looksAt obj1}. This property makes sense to both parties, but in
+the context of the \emph{spygame}, we would like to mainly use immanent
+properties, not volatile like a gaze. More research is required to precise this
+idea of contexts and knowledge class attached to them.
 

### 2013-03-13 paper.tex
Added a couple of macros
Completed the 'Communication' section

@@
 
-\usepackage[utf8x]{inputenc}
+\usepackage[protrusion=true,expansion=true]{microtype}
 \usepackage{cite}
@@
 
+\usepackage{ifthen}
+
+\usepackage[utf8x]{inputenc}
+
 \usepackage[draft]{fixme}
+
+\usepackage{listings}
+\usepackage{alltt}
+\usepackage{fancyvrb}
+
+\newcommand{\concept}[1]{{\small \texttt{#1}}}
+
+\newcommand{\stmt}[1]{{\footnotesize \tt $\langle$ #1\relax$\rangle$}}
+%\newcommand{\stmt}[1]{{\footnotesize $\langle\stmttt#1\relax\rangle$}}
+%\newcommand{\rawstmt}[1]{{\footnotesize \stmttt#1\relax}}
+%\def\stmttt#1 #2 #3\relax{
+%\texttt{\IfBeginWith{#1}{?}{\textbf{#1}}{#1} \IfBeginWith{#2}{?}{\textbf{#2}}{\emph{#2}} \ifthenelse {\equal{#3}{true} \OR \equal{#3}{false}}{\emph{#3}}{\IfBeginWith{#3}{?}{\textbf{#3}} {#3}}}}
+
+%\newcommand{\setstmt}[1]{{\footnotesize [\setstmttt#1\relax]}}
+%\def\setstmttt#1,#2\relax{\stmttt#1\relax, \stmttt#2\relax}
 
@@
 
-\cite{Lemaignan2011a}
+One of the field of human-robot interaction where the introduction of the
+semantic layer has been most beneficial is natural language processing.
+\cite{Lemaignan2011a} details the techniques and the tool called
+\texttt{dialogs} that we have developed for English language parsing and
+grounding, for (minimalist) dialogue management and for verbalisation.
+
+Natural language input (in experiments, we rely on an Android-based interface
+running on a tablet, with Google-based speech recognition, for the input) is
+parsed into a grammatical structure, and each sentence atoms are resolved with
+the help of the ontology to ground classes (when a user says ``pick the can'',
+to which instance of \emph{can} the user refers to?) and actions. Sentences are
+sorted into questions, desires and statements, and processed accordingly.
+
+The system supports quantification (``give me \{a|the|some|all|any|...\} can''),
+thematic roles attached to actions, interactive disambiguation (the robot asks
+questions when needed), anaphora resolution (``give \emph{it} to me'') based on
+dialogue history. It also permits knowledge extension by learning new semantic
+structures (for instance, a sentence like ``learn that cats are animals'' is
+converted into \stmt{Cat subClassOf Animal}, even if \concept{Cat} is not known
+beforehand) and tries to translate \emph{states} (``I'm tired'') into
+\emph{experiences} (\stmt{HUMAN experiences state\_1, state\_1 hasFeature
+tired}).
 
@@
 
-One protoypical example of semantic disambiguation has been given in
+One prototypical example of semantic disambiguation has been given in
 \cite{Ros2010b} with the child's \emph{spygame}: two players are facing
-eachother with a set of random objects inbetween, one player mentally choose
+each other with a set of random objects in-between, one player mentally choose
 one object, and the other player has to guess the object by asking closed
@@
 \subsection{Multi-modal communication}
+
+Because the main communication channel between component is semantic data,
+different communication modalities (explicit like verbal, dectic or based on
+gaze, or implicit like postures) are accessed in an homogeneous way.
+
+The grounding process copes with and integrates the different physical
+modalities at two disctint levels.
+
+First, the grounding module explicitely check for the presence and value of
+particular facts: for instance, when a set of instances match a category (the
+human says ``give me the bottle'' and the robot knows about three bottles), the
+module may decide (it actually depends on the quantifier preceding the class:
+\emph{the}, \emph{a}, \emph{any}...) to discard some of them based on their
+\emph{visibility} by the speaker (implicit communication context built from the
+human posture).
+
+Another example, when the human says ``this'', the robot checks if the human is
+currently both pointing and looking at some object. In that case, \emph{this}
+is replaced by the object focused on.
+
+Note that, while the system benefits from the complementary modalities, they
+are not required. The dialogue system would work as well in a purely verbal
+setting.
+
+The second level of integration of multi-modality is implicit: by computing
+symbolic properties from the geometry, richer descriptions and hence
+discrimination possibilities are available: for instance, if reachability is
+available, the robot may asks ``do you mean the bottle that is accessible to
+me?'' to discriminate between the three bottles. That way, procedures relying
+on discrimination transparently benefit from added modalities.
 

### 2013-03-17 paper.tex
First pass on knowledge model

@@
 
-\cite{Lemaignan2010}
+In our architecture, knowledge manipulation relies on a \emph{semantic
+blackboard}: one central server (the {\sc Oro} server,~\cite{Lemaignan2010})
+stores knowledge as it is produced by the various modules. It also exposes an
+API to query the knowledge base.
+
+Knowledge is represented as RDF triple and using the OWL sub-language. Each
+time triples are added or removed from the knowledge base, an Description
+Logics reasoner (Pellet\footnote{\url{http://clarkparsia.com/pellet/}})
+classifies the whole ontology and appends all possible supplementary
+inferences.
+
+Relying on RDF triples and Description Logics has major advantages like
+the availability of numerous open-source libraries to manipulate the ontology, interoperability with several major on-line knowledge bases (like {\sc OpenCyc}, {\sc WordNet} or {\sc
+DBPedia}), and the formal guarantee of decidability (it is always possible to classify a Description Logics ontology).
+
+It has also notable limitations: RDF triples mean only binary predicates
+(\stmt{subject predicate object}), which constrains the expressiveness of
+the system (or leads to cumbersome reifications). Alternative exists (like {\sc
+KnowRob}~\cite{Tenorth2009a}) that mixes RDF with more expressive logic
+languages like {\sc Prolog} (at the price, however, of other limitations, like
+\emph{closed world assumption} or immutable T-Box). The classification
+performance is another issue: from our experience, with an ontology sized for a
+standard experiment (about 200 instances and 100 classes), classification would
+take about 100ms, which sometimes becomes problematic during interactions.
+Besides, the performances are difficult to predict, since a seemingly
+inoffensive new statement may indirectly change radically the logical
+properties of the whole knowledge model and suddenly lead to much slower
+classification.
 

### 2013-03-17 paper.tex
First pass on theory of mind

@@
 
-Second remark, after some research (see appendix B of~\cite{ma these} for a
+Second remark, after some research (see appendix B of~\cite{Lemaignan2012a} for a
 detailled discussion)  we have decided not to represent neither tasks nor plans
@@
 
-\cite{Warnier2012a}
+Theory of Mind (originally defined in~\cite{Premack1978}) is the
+cognitive ability that a subject has to represent the mental state of another
+agent, possibly including knowledge that contradicts the subject own model: for
+example, a book can be at the same time visible by myself, and not visible by
+you.
+
+This skill, which is essential to understand others' perspectives during
+interactions, appears in children around their third year. It supposes the ability to build, store and retrieve separate models of the knowledge of the interactors.
+
+Our knowledge base implements such a mechanism: when the robots infers a new
+agent has been introduced in the knowledge base, it initializes a new,
+independant, ontology for this agent. All the ontologies that are created share
+the same common-sense knowledge, but rely on each agent's perspective for the
+actual instantiation: the robot (geometrically) computes that the book is in
+its own field of view, but not in the human one. The robot knowledge contains
+the fact \stmt{book isVisible true} while the human model contains \stmt{book
+isVisible false}.
+
+The \cite{Warnier2012a}
 

### 2013-03-21 paper.tex
Added cognitive laod figure
Large proofreading + section on Theory of Mind

--- 2013-03-13 paper.tex
@@
 \usepackage[utf8x]{inputenc}
+\usepackage[T1]{fontenc}
 
@@
 
-Natural interaction and cooperation are the current (dare we say,
-\emph{short-term}) targets for the human-robot interaction community. They are
+Natural interaction and cooperation are on the (dare we say,
+\emph{short-term}) agenda for the human-robot interaction community. They are
 keys to the broad class of \emph{interactive manipulation problems}: several
@@
 cooperation to be successfully achieved. This class of problems involves both
-dialogue and manipulation and they are often not completely defined at
-start-up: they require iterative and interactive resolution.
+dialogue and manipulation and they are often initially underspecified: they
+require iterative and interactive resolution.
 
@@
 prerequisites of these challenges, and giving them experimental reality on the
-robots: what is required for a sentence like ``Let's dress the table together''
+robots: what is required for sentences like ``Let's dress the table together''
 to be understood by the robot, correctly interpreted in the spatial and
 temporal context of the interaction, and eventually transformed into a set of
-actions?
+actions.
 
@@
 
-We focus in this paper on that last item, to present what that means, and what
-are the benefits and drawbacks.
+The last item is the focus of this paper: what does ``knowledge-centric''
+architecture means, and what are its benefits and drawbacks.
 
@@
 
-While we do not present them here, the ideas and techniques pertaining to this
-paper have been all implemented and tested on several robots (see previously
+While we do not present experiments in this paper, the ideas and techniques we
+present all have been implemented and tested on several robots (see previously
 reported experiments in~\cite{Lemaignan2010, Ros2010b, Lemaignan2011a,
-Warnier2012a}). Videos presenting some of these experimental results can also
+Warnier2012a}). Also, videos presenting some of these experimental results can
 be watched online at~\url{http://www.laas.fr/~slemaign}.
@@
 
-We first describe with more details how the deliberative layer of our robots is
-built and why this monolithic functional group enables what we call a
-\emph{knowledge-oriented} architecture. The first section also introduce our
-central knowledge management tool, the ORO server.
+In the next section, we first introduce the deliberative layer of our robots at
+the functional level. It represents what we call a \emph{knowledge-oriented}
+architecture, built around a central knowledge management tool called the {\sc
+Oro} server.
 
-The next sections go over each of the facets of this deliberative layer: how
-knowledge is produced asynchronously from geometric reasoning, how the robot
-conducts grounded multi-modal interaction with humans, how knowledge is used by
-decision-making components like the robot controller or the task planner, and
-finally, how our knowledge-centric architecture enables the implementation of
-specific internal cognitive processes.
+The following sections go over each of the facets of this deliberative layer:
+how knowledge is produced asynchronously from geometric reasoning, how the
+robot conducts grounded multi-modal interaction with humans, how knowledge is
+used by decision-making components like the robot controller or the task
+planner, and finally, how our knowledge-centric architecture enables the
+implementation of specific internal cognitive processes.
 
-We conclude the article with a discussion about the strenghts and weaknesses of
-such a knowledge-oriented architecture, fron three perspectives: the architect,
-the IA expert and the cognitician.
+We conclude the article with a discussion about the strengths and weaknesses of
+such a knowledge-oriented architecture, from three perspectives: the software
+architect, the IA expert and the cognitician.
 
@@
 \cite{Alami2011}
+\ref{fig|archi}
 
@@
         \includegraphics[width=1.7\columnwidth]{archi}
-        \caption{Overview of the LAAS deliberative layer.}
+        \caption{Overview of the LAAS deliberative layer. Knowledge is
+        centrally managed in an active \emph{semantic blackboard}, pictured
+        above with a thick border.}
         \label{fig|archi}
 \end{figure*}
-
-We can give a broader look at the knowledge and the streams of knowledge in our
-systems.  Based on the experience gained while developing and deploying {\sc
-ORO}, our ontology-based knowledge server, we have presented how symbolic
-knowledge could be produced from perception and geometric reasoning in modules
-like {\sc SPARK}, a grounded, perspective-aware, geometric reasoner. We have
-seen how symbolic knowledge could be reused by different control systems and
-task planners like {\sc Cram}, {\sc SHARY}, {\sc pyRobots}, the {\sc CLSU
-Toolkit} or {\sc HATP} and how they take advantage of semantic abstractions
-provided by knowledge base. We have also presented the bidirectional
-integration of {\sc Dialogs}, a natural language processor for English, with
-the knowledge base.
-
-Altogether, these components compose an architecture that we call
-\emph{knowledge-oriented}:
-
-\begin{itemize}
-
-    \item{Knowledge is explicitly stored in one central and consistent
-    repository of facts, accessible by all modules.}
-
-    \item{Knowledge is represented in a strict formalism (OWL statements) and
-    with a clearly defined vocabulary (stated in the common-sense ontology).}
-
-    \item{The first two points enable both a loosely-coupled architecture where
-    modules can very easily be removed or replaced by other ones as long as
-    they share the same semantics (modules are defined by the knowledge they
-    produce),}
-
-    \item{and a \emph{symbolic} reactive, event-driven approach to supervision.
-    By managing events at the same level as the reasoner, we take full
-    advantage of the inference abilities of ORO to trigger events whose
-    \texttt{true} conditions can be inferred.}
-
-    \item{Finally, this architecture allows for the combination of very
-    different knowledge modalities in a single homogeneous environment,
-    bringing mutual benefits to components. For instance, the dialogue
-    processing module can perfectly run without any geometric
-    perception, but its disambiguation routines can transparently
-    benefit from it when available (since richer symbolic descriptions of
-    objects are then available).}
-
-\end{itemize}
 
@@
 and \textsc{Dialogs} constantly build and push new symbolic contents about the
-world to ORO where it becomes accessible to decisional layers. In parallel, ORO
+world to {\sc Oro} where it becomes accessible to decisional layers. In parallel, {\sc Oro}
 relies on reasoning in a \textit{top-down} way to produce new facts that may
 trigger in return physical behaviours.
+
+\subsection{Knowledge model}
+
+In our architecture (Fig.~\ref{fig|archi}), knowledge manipulation relies on a
+\emph{semantic blackboard}: one central server (the {\sc Oro}
+server~\cite{Lemaignan2010}) stores knowledge as it is produced by each of the
+deliberative components. It conversely exposes an API to query the knowledge
+base.
+
+Knowledge is represented as RDF triples in the OWL sub-language. Each
+time triples are added or removed from the knowledge base, a Description
+Logics reasoner ({\sc Pellet}\footnote{\url{http://clarkparsia.com/pellet/}})
+classifies the whole ontology and inserts all possible inferred triples.
+
+Relying on RDF triples and Description Logics has major advantages like the
+availability of numerous open-source libraries to manipulate the ontology,
+interoperability with several major on-line knowledge bases (like {\sc
+OpenCyc}, {\sc WordNet} or {\sc DBPedia}), open-world reasoning, and the formal
+guarantee of decidability (it is always possible to classify a Description
+Logics ontology).
+
+It has also notable limitations: RDF triples imply only binary predicates
+(\stmt{subject predicate object}), which constrains the expressiveness of the
+system or leads to cumbersome reifications. Alternative exists (like {\sc
+KnowRob}~\cite{Tenorth2009a}) that mixes RDF with more expressive logic
+languages like {\sc Prolog}, at the price, however, of other limitations, like
+closed-world reasoning or immutable T-Box. The classification
+performance is another issue: from our experience, with an ontology sized for a
+standard experiment (about 100 classes and 200 instances), classification
+typically take about 100ms, which becomes problematic during interactions.
+Besides, the performances are difficult to predict, since a seemingly
+inoffensive new statement may indirectly change radically the logical
+complexity of the whole knowledge model and lead to notable degradation of
+classification time.
+
+This knowledge model also largely exclude representation of continuous
+phenomena (like time) or uncertain phenomena. When required, these have to be
+managed by dedicated components.
+
+Amongst the features of {\sc Oro}, the system can dynamically creates
+new independent knowledge models (\ie independent instantiations of the
+ontology). This feature is used for instance to store separately the beliefs of
+each of the agents the robot interacts with (see section~\ref{sect|tom}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Situation Assessment}
+\label{sect|sit-ass}
+
+\begin{figure}
+        \centering
+        \includegraphics[width=\columnwidth]{spark-oro}
+    \caption{Functional overview of knowledge base ({\sc Oro} server, top part) and the geometric situation assessment module (\emph{SPARK}, bottom part)}
+        \label{fig|spark-oro}
+\end{figure}
+
+\cite{Sisbot2011}
+
+\subsection{Ego-centric and allo-centric frames}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Communication}
+\label{sect|com}
+
+\subsection{Natural language grounding}
+
+Natural language processing is one of the fields of human-robot interaction
+where the introduction of the semantic layer has been most beneficial.
+We detail in~\cite{Lemaignan2011a} the techniques and the tool called
+\texttt{dialogs} that we have developed for English language parsing and
+grounding, along with verbalisation and (minimalist) dialogue management.
+
+Natural language input (in experiments, we rely on an Android-based interface,
+with Google speech recognition) is parsed into a grammatical structure, and
+each sentence atoms are resolved with the help of the ontology to ground
+concepts like objects (\ie when a user says ``pick the can'', resolve to which
+instance of \emph{can} the user is referring to) and actions. Sentences are
+sorted into questions, desires and statements, and processed accordingly.
+
+The system supports quantification (``give me \{a | the | some | all | any | ...\} can''),
+thematic roles (action-specific predicates that qualify the actions), interactive disambiguation (the robot asks
+questions when it needs more information), anaphora resolution (``give \emph{it} to me'') based on
+dialogue history. It also permits knowledge extension by learning new semantic
+structures (for instance, a sentence like ``learn that cats are animals'' is
+converted into \stmt{Cat subClassOf Animal}, interprets common temporal and place adverbs (like \emph{above} or \emph{tomorrow}) and translates to a certain extend \emph{states} (``I'm tired'') into
+\emph{experiences} (\stmt{HUMAN experiences state\_1, state\_1 hasFeature
+tired}).
+
+\subsection{Limits of disambiguation at semantic level}
+
+One prototypical example of semantic disambiguation has been given in
+\cite{Ros2010b} with the child game \emph{spygame}: two players are facing
+each other with a set of random objects in-between, one player mentally choose
+one object, and the other player has to guess the object by asking closed
+questions like \emph{Is your object small or large?} Based on the knowledge it
+has acquired, the robot is able to minimize the number of questions required to
+find the object.
+
+When playing this kind of game, however, the issue arises that the robot has no
+way to select which knowledge about the object is relevant in the interaction
+context. For instance, the knowledge base may store facts like \stmt{obj1 type
+ActiveConcept} (which internally means that this concept was mentioned in a
+discussion in the last few seconds): this information is not a relevant
+property of \concept{obj1} when trying to disambiguate concepts with humans.
+This distinction between \emph{internal knowledge} (meaningful to
+the system only) and \emph{common knowledge} (whose meaning is understood by
+all the interactors) has not been properly dealt with in our architecture.
+
+Besides, even knowledge that belongs the \emph{common knowledge} may not be
+appropriate in a given interaction context. For instance, the system may
+compute that at a given instant the human is looking at the object: \stmt{human
+looksAt obj1}. This property makes sense to both parties, but in the context of
+the \emph{spygame}, we would like to mainly use immanent properties, not
+volatile like a gaze. More research is required to indentify relevant
+interaction contexts and knowledge classes attached to them.
+
+\subsection{Multi-modal communication}
+
+Because all components rely on the same RDF formalism to format their outputs,
+the different communication modalities (\emph{explicit} like verbal, deictic or
+based on gaze, or \emph{implicit} like postures) are presented in an
+homogeneous way.
+
+The dialogue grounding process makes use of them at two distinct levels.
+
+First, particular steps of the grounding process explicitly check for the
+presence and value of specific facts: for instance, when a set of instances
+match a category (the human says ``give me the bottle'' and the robot knows
+about three bottles), the module may decide (it actually depends on the
+quantifier preceding the class) to discard some of them based on their
+\emph{visibility} by the speaker (implicit communication context built on the
+human posture).
+
+Another example, when the human says ``this'', the robot checks if the human is
+currently pointing at some object. In that case, \emph{this} is replaced by the
+object focused on.
+
+Note that, while the system benefits from the complementary modalities, they
+are not all required. The dialogue system would for instance run equally well
+with only the verbal modality.
+
+The second level of integration of multi-modality is implicit: by computing
+symbolic properties from the geometry, richer descriptions and hence
+discrimination possibilities are available: for instance, if reachability is
+available, the robot may asks ``do you mean the bottle that is accessible to
+me?'' to discriminate between the three bottles. That way, procedures relying
+on discrimination transparently benefit from added modalities.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Robot control}
+\label{sect|ctrl}
+
+\subsection{Modeling the knowledge related to the interaction}
+
+We split the interaction situations stemming from the situation assessment and
+communication components in two categories: \emph{desires} and
+\emph{experiences}.
+
+Experiences comprise of emotions and questions
+
+\subsubsection{Event-driven control}
+
+\subsection{Task planning}
+
+In order to devise how a given goal can be accomplished, the robot has to
+elaborate a plan, \ie a set of actions to be achieved by the robot and its
+human partner(s).  We use in our architecture the HATP planner (for Human Aware
+Task Planner~\cite{Alili2009}).  HATP is based on a Hierarchical Task Network
+(HTN) refinement, which performs an iterative task decomposition into sub-tasks
+until reaching atomic actions.  The \emph{planning domain} defines the set of
+methods describing how to decompose a task and can be seen as the procedural
+knowledge of the robot.
+
+In order to produce a collaborative behaviour, the robot plans not only for
+itself but also for the other agents. The planning domain of each agent is
+instantiated from the agent-specific model in the {\sc Oro} server. The resulting
+plan, called \emph{shared plan} is a set of action streams, one per
+agent involved in the goal achievement. HATP can also be tuned by setting up
+different costs depending on the actions to apply and by taking into account a
+set of constraints called \emph{social rules}. This tuning aims at adapting the
+robot's behaviour according to the desired level of cooperation of the robot.
+
+Two important remarks: because HATP is a generic symbolic task planner, we have
+been able to design a planning domain at a semantic level which is close to the
+one used in the human-robot dialogue (the planner vocabulary contains concepts
+like \texttt{give}, \texttt{table}, \texttt{is on}...). Hence only a few
+ontology rules have been required to map both the knowledge extracted from the
+situation assessment and the statements originated from the verbal interaction
+to the planner domain.
+
+Second remark, after some research (see appendix B of~\cite{Lemaignan2012a} for
+a detailed discussion)  we have decided not to represent neither the planning
+domain nor the resulting plans in the knowledge base: the planning domain (with
+tasks pre- and postconditions) is stored in a specific format, outside of the
+knowledge repository, and the plans are directly communicated to the robot
+controller. From our experience, the OWL formalism is not a satisfying
+representation language for concepts that are both tightly bound to time
+representation and non-monotonic.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Internal cognitive processes}
+\label{sect|intern}
+
+\subsection{Cognitive load}
+
+\begin{figure}
+        \centering
+        \includegraphics[width=\columnwidth]{figs/cognitive_load.png}
+        \caption{This diagram shows the average number of interactions with the
+        knowledge base during a one-hour long experiment with the robot.
+        Interactions are either knowledge alteration (addition/removal of
+        statements) or queries. They are averaged on a sliding window of 60 seconds.}
+        \label{fig|cognitiveload}
+\end{figure}
+
+Fig.~\ref{fig|cognitiveload} is a plot of the log of interactions between the
+robot modules and the knowledge base during a one-hour long experiment. Because
+most of the communication between modules goes in fact through the knowledge
+base in our architecture (Fig.~\ref{fig|archi}), it actually pictures the
+intensity of internal information exchanges between the different deliberative
+component of the robot over time. Since communication actually relies on
+grounded semantic information, we can call this level of interaction the
+\emph{cognitive load} of the robot.
+
+This simplistic measurement needs to be refined to take into account the fact
+that the raw number of RPC calls to the knowledge base only imperfectly
+reflects the actual cognitive processes running on, but it opens interesting
+perspective nevertheless. It would be interesting for example to compare
+and look for correlations with the cognitive load of a human partner (measured
+via EEG of qualitative feedback).
+
+\subsection{Theory of Mind}
+\label{sect|tom}
+
+Theory of Mind (originally defined in~\cite{Premack1978}) is the cognitive
+ability that a subject possesses to represent the mental state of another
+agent, possibly including knowledge that contradicts the subject own model: for
+example, a book can be at the same time \emph{visible} by myself, and \emph{not
+visible} by you.
+
+This skill, which is essential to understand others' perspectives during
+interactions, appears in children around their third year. It supposes the ability to build, store and retrieve separate models of the knowledge of the interactors.
+
+Our knowledge base implements such a mechanism: when the robots infers a new
+agent has been introduced in the knowledge base, it initializes a new,
+independent, ontology for this agent. All the ontologies that are created share
+the same common-sense knowledge, but rely on each agent's perspective for the
+actual instantiation: the robot (geometrically) computes that the book is in
+its own field of view, but not in the human one. The robot knowledge contains
+the fact \stmt{book isVisible true} while the human model contains \stmt{book
+isVisible false}.
+
+The \cite{Warnier2012a}
+
+\subsection{Working memory}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+\label{sect|conclusion}
+
+\subsection{Knowledge-oriented architecture}
+
+Altogether, the components we have presented compose an architecture that we
+call \emph{knowledge-oriented}:
+
+\begin{itemize}
+
+    \item{Knowledge is explicitly stored in one central and consistent
+    repository of facts, accessible by all modules.}
+
+    \item{Knowledge is represented in a strict formalism (OWL statements) and
+    with a clearly defined vocabulary (stated in the common-sense ontology).}
+
+    \item{The first two points enable both a loosely-coupled architecture where
+    modules can very easily be removed or replaced by other ones as long as
+    they share the same semantics (modules are defined by the knowledge they
+    produce),}
+
+    \item{and a \emph{symbolic}, reactive, event-driven approach to robot control.
+    By managing events at the same level as the reasoner, we take full
+    advantage of the inference abilities of {\sc Oro} to trigger events whose
+    \texttt{true} conditions can be inferred.}
+
+    \item{Finally, this architecture allows for the combination of very
+    different knowledge modalities in a single homogeneous environment,
+    bringing mutual benefits to components. For instance, the dialogue
+    processing module can perfectly run without any geometric
+    perception, but its disambiguation routines can transparently
+    benefit from it when available (since richer symbolic descriptions of
+    objects are then available).}
+
+\end{itemize}
 
@@
 
-This architecture may also contribute to bring closer robotics and psychology:
-it provides clear entry points to implement some classical psychology tests to
-robots. For instance, we presented experiments focused on issues related to
-perspective taking. By explicitly enabling independent modeling of the beliefs
-of each agent, our architecture is especially well suited to set up cognitive
-and psychological experiments that involve a theory of mind, such as
-\emph{False-Belief} experiments, as recently presented in~\cite{Warnier2012a}.
-
-\subsection{Knowledge Model}
-
-\cite{Lemaignan2010}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Situation Assessment}
-\label{sect|sit-ass}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=\columnwidth]{spark-oro}
-        \caption{Functional overview of knowledge base (\emph{oro-server}, top part) and the geometric situation assessment module (\emph{SPARK}, bottom part)}
-        \label{fig|spark-oro}
-\end{figure}
-
-\cite{Sisbot2011}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Communication}
-\label{sect|com}
-
-\subsection{Natural language grounding}
-
-One of the field of human-robot interaction where the introduction of the
-semantic layer has been most beneficial is natural language processing.
-\cite{Lemaignan2011a} details the techniques and the tool called
-\texttt{dialogs} that we have developed for English language parsing and
-grounding, for (minimalist) dialogue management and for verbalisation.
-
-Natural language input (in experiments, we rely on an Android-based interface
-running on a tablet, with Google-based speech recognition, for the input) is
-parsed into a grammatical structure, and each sentence atoms are resolved with
-the help of the ontology to ground classes (when a user says ``pick the can'',
-to which instance of \emph{can} the user refers to?) and actions. Sentences are
-sorted into questions, desires and statements, and processed accordingly.
-
-The system supports quantification (``give me \{a|the|some|all|any|...\} can''),
-thematic roles attached to actions, interactive disambiguation (the robot asks
-questions when needed), anaphora resolution (``give \emph{it} to me'') based on
-dialogue history. It also permits knowledge extension by learning new semantic
-structures (for instance, a sentence like ``learn that cats are animals'' is
-converted into \stmt{Cat subClassOf Animal}, even if \concept{Cat} is not known
-beforehand) and tries to translate \emph{states} (``I'm tired'') into
-\emph{experiences} (\stmt{HUMAN experiences state\_1, state\_1 hasFeature
-tired}).
-
-\subsection{Disambiguation at semantic level}
-
-One prototypical example of semantic disambiguation has been given in
-\cite{Ros2010b} with the child's \emph{spygame}: two players are facing
-each other with a set of random objects in-between, one player mentally choose
-one object, and the other player has to guess the object by asking closed
-questions like \emph{Is your object small or large?}. Based on the knowledge it
-has acquired, the robot is able to minimize the number of questions required to
-find the object.
-
-When playing this kind of game, however, the issue arises that the robot has no
-way to select which knowledge about the object is relevant in the interaction
-context. For instance, the knowledge base may store facts like \stmt{obj1 type
-ActiveConcept} (which internally means that this concept was mentioned in a
-discussion in the last few seconds): this information is not a relevant
-property of \concept{obj1} when trying to disambiguate concepts with humans.
-This distinction between \emph{internal knowledge} (meaningful to
-the systme only) and \emph{common knowledge} (whose meaning is understood by
-all the interactors) has not been properly dealt with in our architecture.
-
-Besides, even knowledge that belongs to what we call \emph{common knowledge},
-may not be appropriate in a given interaction context. For instance, the system
-may compute that at a given instant, the human is looking at the object:
-\stmt{human looksAt obj1}. This property makes sense to both parties, but in
-the context of the \emph{spygame}, we would like to mainly use immanent
-properties, not volatile like a gaze. More research is required to precise this
-idea of contexts and knowledge class attached to them.
-
-\fxnote{Mention failures like 'placeOing'}
-
-\subsection{Multi-modal communication}
-
-Because the main communication channel between component is semantic data,
-different communication modalities (explicit like verbal, dectic or based on
-gaze, or implicit like postures) are accessed in an homogeneous way.
-
-The grounding process copes with and integrates the different physical
-modalities at two disctint levels.
-
-First, the grounding module explicitely check for the presence and value of
-particular facts: for instance, when a set of instances match a category (the
-human says ``give me the bottle'' and the robot knows about three bottles), the
-module may decide (it actually depends on the quantifier preceding the class:
-\emph{the}, \emph{a}, \emph{any}...) to discard some of them based on their
-\emph{visibility} by the speaker (implicit communication context built from the
-human posture).
-
-Another example, when the human says ``this'', the robot checks if the human is
-currently both pointing and looking at some object. In that case, \emph{this}
-is replaced by the object focused on.
-
-Note that, while the system benefits from the complementary modalities, they
-are not required. The dialogue system would work as well in a purely verbal
-setting.
-
-The second level of integration of multi-modality is implicit: by computing
-symbolic properties from the geometry, richer descriptions and hence
-discrimination possibilities are available: for instance, if reachability is
-available, the robot may asks ``do you mean the bottle that is accessible to
-me?'' to discriminate between the three bottles. That way, procedures relying
-on discrimination transparently benefit from added modalities.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Robot Control}
-\label{sect|ctrl}
-
-\subsection{Modeling the knowledge related to the interaction}
-
-We split the interaction situations stemming from the situation assessment and
-communication components in two categories: \emph{desires} and
-\emph{experiences}.
-
-Experiences comprise of emotions and questions
-
-\subsection{Task planning}
-
-In order to devise how a given goal can be accomplished, the robot has to
-elaborate a plan,\ie a set of actions to be achieved by the robot and its human
-partners.  We use in our architecture the HATP planner~\cite{Alili2009} (for
-Human Aware Task Planner).  HATP is based on a Hierarchical Task Network (HTN)
-refinement, which performs an iterative task decomposition into sub-tasks until
-reaching atomic actions.  The planning domain defines a set of methods
-describing how to decompose a task and can be seen as the procedural knowledge
-of the robot.
-
-In order to produce a collaborative behaviour, the robot plans not only for
-itself but also for the other agents. The planning domain of each agent is
-instantiated from the agent-specific model in the ORO server. The resulting
-plan, called ``shared plan'' is a set of actions that form a stream for each
-agent involved in the goal achievement. HATP can also be tuned by setting up
-different costs depending on the actions to apply and by taking into account a
-set of constraints called \emph{social rules}. This tuning aims at adapting the
-robot's behaviour according to the desired level of cooperation of the robot.
-
-Two important remarks: because HATP is a generic symbolic task planner, we have
-been able to design a planning domain at a semantic level which is close to the
-one used in the human-robot dialogue (the planner vocabulary contains concepts
-like \texttt{give}, \texttt{table}, \texttt{is on}...). Hence only a few
-ontology rules have been required to map both the knowledge extracted from the
-situation assessment and the statements originated from the verbal interaction
-to the planner domain.
-
-Second remark, after some research (see appendix B of~\cite{ma these} for a
-detailled discussion)  we have decided not to represent neither tasks nor plans
-in the ontology: the planning domain (with tasks pre- and postconditions) is
-stored in a specific format, outside of the knowledge repository, and the
-resulting plans are directly communicated to the robot controller. The ontology
-formalism that we have adopted (namely OWL) does not appear to be a satisfaying
-reprepresentation language for concepts that are both tightly bound to time
-representation and non-monotonic. Other researches (\cite{tenorth...}) have attempted with more success (\fixme{regarder ca!!}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Internal Cognitive Processes}
-\label{sect|intern}
-
-\subsection{Theory of Mind}
-
-\cite{Warnier2012a}
-
-\subsection{Working Memory}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-\label{sect|conclusion}
+\subsection{Three perspectives on knowledge manipulation for robots}
 
@@
 the \emph{logician perspective}: does the promise of automated reasoning for
-robots really comes to fruition with ontologies? And eventually the
+robots really comes to fruition with ontologies? And finally the
 \emph{cognitician perspective} or how explicit knowledge manipulation endows
@@
 
-\subsection{The Architect view: loose coupling and modalities merging}
+\subsubsection{The Architect view: loose coupling and modalities merging}
 
@@
 Since these statements are anchored in the robot's pool of knowledge, they
-actually convey \emph{meaning} through the system (for instance, when the
-dialog module or the task planner or the execution controller uses the symbol
-{\tt Give} \emph{means}, it refers to the abstract class of all give actions, which are also stated
-to be a kind of purposeful actions). This has one important consequence from
-the perspective of the system architecture: modules interfaces are only defined
-in term of knowledge consumption and/or production. This brings good decoupling
-properties. For instance, the output of the dialog module is a new
-\emph{situation desired} by the speaker. This module can be transparently
-replaced or completed by any other module that produces the same chuncks of
-knowledge. Only a thin protocol layer remains to get the modules to interact
-together. This allows us to implement (otherwise difficult) merging of
-radically different perception modalities like vision (the robot sees an object on
-a table) and verbal description (a human says to the robot another object is
-under the table): both are converted to RDF statements ({\tt obj1 type Object},
-{\tt obj1 isOn table1} and {\tt obj2 type Object}, {\tt obj2 isUnder table1})
-and made available to the other robot modules.
+actually convey shared and unambiguous \emph{meaning} through the system. This
+has one important consequence from the perspective of the system architecture:
+modules interfaces are only defined in term of knowledge consumption and/or
+production. This brings good decoupling properties. For instance, the output of
+the dialog module is a new \emph{situation desired} by the speaker. This module
+can be transparently replaced or completed by any other module that produces
+the same chuncks of knowledge. Only a thin protocol layer remains to get the
+modules to interact together. This allows us to implement (otherwise difficult)
+merging of radically different perception modalities like vision (the robot
+sees an object on a table: the output would look like \stmt{obj1 isOn table1})
+and verbal description (a human says to the robot the object on the table
+is red: it leads to a query \stmt{? isOn table1} and then a new assertion:
+\stmt{obj1 hasColor red}).
 
@@
 
+On the sensing side also we have in certain cases to bypass the semantic layer
+(extension of the symbolic task planner with geometric constraints is one
+example).
+
 And obviously, many aspects like uncertainty management, modal logics,
-non-monotonic reasoning,... remain to be explored to gain the level of
-expressiveness required to cover many of the more complex human-robot
-interaction scenarii.
+non-monotonic reasoning,... are not well addressed with the Description Logic
+formalism, and remain to be explored to gain the level of expressiveness
+required to cover many of the more complex human-robot interaction scenarii.
 
-\subsection{The Logician view: the importance of trivial inferences}
+\subsubsection{The Logician view: the importance of trivial inferences}
 
@@
 simple yet non-trivial reasoning. As a matter of fact, only very few of such
-reasoning cases where positively identified in our scenarii and experiments
-(and consequently implemented as rules in ORO).
+prototypical reasoning cases where positively identified in our scenarii and experiments
+(and consequently implemented as rules in {\sc Oro}).
 
 The design choices of our architecture partially explain that fact: first, the
-planning task (which is the prototypical reasoning task) is delegated to a
-dedicated, external planner. Then, time is not represented in ORO, and
+planning task (which is a typical reasoning task) is delegated to a
+dedicated, external planner (HATP). Then, time is not represented in {\sc Oro}, and
 consequently no temporal reasoning takes place at this level: action
@@
 semantics would be desirable to better stress the expressiveness and inference
-abilities provided by description logics.
+abilities provided by Description Logics.
 
@@
 constraints, transitivity, etc.) and encode a large amount of common-sense
-knowledge that would be tedious, to say the least, to manually assert. These
-trivial inferences are all the more important that an expressive knowledge
-representation language is used: when a language like OWL allows to directly
-represent high-level concepts like partitions, cardinality restrictions,
-properties' ranges and domains, it leads to a more implicit (because more
-abstract) description of the vocabulary that in turn requires more underlying
-reasoning. With the progress in the understanding of the relations between
-expressiveness and (tractable) satisfiability, along with the progress of
-reasoners, more and more of the inferences do not need to be explicit anymore,
-and consequently move behind the scenes.
+knowledge that can not be otherwise asserted. These trivial inferences are all
+the more important that an expressive knowledge representation language is
+used: when a language like OWL allows to directly represent high-level concepts
+like partitions, cardinality restrictions, properties' ranges and domains, it
+leads to a more implicit (because more abstract) description of the vocabulary
+that in turn requires more underlying reasoning. With the progress in the
+understanding of the relations between expressiveness and (tractable)
+satisfiability, along with the progress of reasoners, more and more of the
+inferences do not need to be explicit anymore, and consequently are delegated
+to the reasoner.
 
-And we predict that \emph{common-sense encoding} is likely to remain the main
-application of reasoning in our robotic architectures, where reasoning related
+We think that \emph{common-sense encoding} is likely to remain the main
+application of reasoning in our architecture, where reasoning related
 to decision making mostly happen outside the knowledge representation system.
 
-\subsection{The Cognician view: palpable knowledge and semantic thinking}
+\subsubsection{The Cognician view: palpable knowledge and semantic thinking}
 
-The main motive goal was to transform the knowledge in the robot from some ubiquitous,
+The original motivation to introduce explicit knowledge management in our
+architecture was to transform the knowledge in the robot from some ubiquitous,
 pervasive, multi-modal and, most importantly, mostly undefined feature of the
-system into an observable, quantifiable, manipulable resource, what we could
-call a \emph{palpable} feature.
+system into an observable, quantifiable, manipulable resource: a
+\emph{palpable} feature of the system.
 
-This transformation, both from the technical point of view (the ORO server, the
-ontologies, the bindings, etc.), and as a more subtle change in the practises
-related to the development of robotic components, is the main contribution of
-this thesis.
+This transformation, both from the technical point of view (the {\sc Oro}
+server, the ontologies, the bindings, etc.), and as a more subtle change in the
+practises related to the development of robotic components, is probably the
+main outcome of this work.
 
@@
 compatibility of interfaces. When defining the frontiers of a robotic
-component, we do not think anymore only in term of ``is the interface complete
-and self-contained'', but also in term of ``is the semantic complete and
-consistent?''. This allow a deeper, more correct modularity: two modules that
+component, we do not think anymore only in term of \emph{is the interface complete
+and self-contained}, but also in term of \emph{is the semantic complete and
+consistent?}. This allow a deeper, more correct modularity: two modules that
 share the same, well-defined semantic can be confidently exchanged. When we
@@
 
+Even more, this architecture also contribute to bring closer robotics and
+cognitive psychology: it provides clear entry points to implement some
+classical psychology tests to robots, be it related to perspective taking, to
+language understanding, to a theory of mind (such as the \emph{False-Beliefs}
+experiment), etc.
+
 \section*{Acknowledgment}

### 2013-03-22 paper.tex
spell checking

@@
 This article presents our design choices, the articulations between the diverse
-deliberative components of the robot, and the strenghts and weaknesses of this
+deliberative components of the robot, and the strengths and weaknesses of this
 approach. We show that explicit knowledge management is not only a convenient
@@
 the \emph{spygame}, we would like to mainly use immanent properties, not
-volatile like a gaze. More research is required to indentify relevant
+volatile like a gaze. More research is required to identify relevant
 interaction contexts and knowledge classes attached to them.
@@
 can be transparently replaced or completed by any other module that produces
-the same chuncks of knowledge. Only a thin protocol layer remains to get the
+the same chunks of knowledge. Only a thin protocol layer remains to get the
 modules to interact together. This allows us to implement (otherwise difficult)
@@
 Another consequence is that the communication channels are {\it
-defacto} defined by the semantics of the information they
+de facto} defined by the semantics of the information they
 convey. This allows the system designers to think in term of which module
@@
 
-\subsubsection{The Cognician view: palpable knowledge and semantic thinking}
+\subsubsection{The Cognitician view: palpable knowledge and semantic thinking}
 

### 2013-03-22 paper.tex
Fixed affiliation

@@
 \author{Séverin Lemaignan and Rachid Alami\\
-CNRS-LAAS, 7 av. du Colonel Roche, F-31077 Toulouse, France\\
-Université de Toulouse, UPS, INSA, INP, ISAE, LAAS, F-31077 Toulouse, France\\
-{\tt surname.name@laas.fr}
+CNRS, LAAS, 7 avenue du Colonel Roche, F-31400 Toulouse, France\\
+Univ de Toulouse, LAAS, F-31400 Toulouse, France\\
+{\tt severin.lemaignan@laas.fr}, {\tt rachid.alami@laas.fr}
 }

### 2013-03-22 paper.tex
Fixed some references

@@
 We have chosen to tackle the challenge from several ends: human-aware
-navigation and motion planning (\cite{...Mamoun}), situation assessment coupled
-with motion planning (\cite{...}), projection of ``mightabilities'' that
-anticipates what surrounding agents may do (\cite{Amit...}), and the design and
-deployment of a pervasive \emph{knowledge-centric} software architecture.
+navigation and motion planning (\cite{Mainprice2011}), situation assessment
+coupled with motion planning (\cite{Mainprice2012}), projection of
+``mightabilities'' that anticipates what surrounding agents may do
+(\cite{Pandey2011}), and the design and deployment of a pervasive
+\emph{knowledge-centric} software architecture.
 

### 2013-03-22 paper.tex
Added section on situation assessment

@@
 
-\cite{Sisbot2011}
+Anchoring perceptions in a symbolic model requires perception abilities and
+their symbolic interpretation. We rely on a dedicated geometric and temporal
+reasoning module (as named on Fig.~\ref{fig|archi}) called SPARK (\emph{SPAtial
+Reasoning \& Knowledge}~\cite{Sisbot2011}). It is a situation assessment
+reasoner that generates symbolic knowledge from the geometry of the
+environment with respect to relations between objects, robots and humans,
+also taking into account the different perspective that each agent has on
+the environment (Fig.~\ref{fig|spark-oro}).
+
+SPARK is an \emph{amodal} geometric model of the environment that serves both
+as basis for the fusion of the perception modalities and as bridge with the
+symbolic layer. This geometric model is continuously updated at run-time by the
+robot based on its sensors. In our experiments, objects are identified and
+localised through 2D barcodes.  On the other hand, the human pose is tracked
+with a Kinect-like device (assisted by motion capture to accurately track the
+head motion, which is required to compute what the human is looking at).
+
+\subsection{Computed symbolic relations}
+
+\subsubsection{Agent-independent locations}
+
+Absolute spatial relations (\concept{isOn}, \concept{isIn}, \concept{isNextTo},
+\concept{isAbove}...) are computed for every pairs of object, taking into
+account relative sizes for relations like \concept{isNextTo}. Other
+agent-independent properties like \concept{isMoving} are also computed.
+
+\subsubsection{Agent-dependent placements}
+
+The space surrounding each agent is divided into 16 regions (like
+\emph{back-left}, \emph{front-near}, etc.). Based on this partition, symbolic
+relations between each pair (agent, object) is computed and added to the
+knowledge base.
+
+\subsubsection{Affordances}
+
+SPARK also computes affordances: \concept{sees}, \concept{looksAt},
+\concept{pointsAt}, \concept{canReach}. Visibility (with two levels
+corresponding to the field of view and the field of attention) and pointing are
+computed by placing virtual cameras in the SPARK 3D model, while reachability
+relies on inverse kinematics. Several filters (based on hystersis, expected
+object behaviour, etc.) allow to partially account for perception
+noise~\cite{Warnier2012a}.
 
 \subsection{Ego-centric and allo-centric frames}
+
+SPARK enables \emph{perspective-taking}: spatial relations between entities can
+be computed from different viewpoints, which let the robot build for each
+agents it interacts with a different, \emph{perspective-aware} symbolic model
+of the environment. These models are separately stored in the knowledge base.
+
+This allows us to deal with ambiguities that arises when one speaker refers to
+an object within a reference system (or changes the reference system, \ie
+switches perspective) without making the reference frame
+explicit~\cite{Breazeal2006, Ros2010}.
+
+As a result, the robot stores models on the environment either in the
+\emph{ego-centric} reference frame (from the robot perspective) or in
+addressee-centred (\ie, \emph{allo-centric}) frames.
 

### 2013-03-22 paper.tex
Added section on Working memory and finish Theory of Mind

@@
 
-The \cite{Warnier2012a}
+One classical application of this cognitive skills is the so-called
+\emph{False-Belief} experiment~\cite{Leslie2000}: a child is asked to watch a
+scene where two people, A and B, manipulate objects. Then A leaves and B hides
+away one object. When A comes back, we ask the child ``where do you think A
+will look for the object?''. Before acquiring a theory of mind, children are
+not able to separate their own (true) model of the world (where they know that
+the object was hidden) from the model of A, which contains \emph{false beliefs}
+on the world (A still thinks the object is at its original position since he
+did not see B hiding it). Using seperate knowledge models in the knowledge
+base, we have been able to replicate this experience on our
+robots~\cite{Warnier2012a}.
 
 \subsection{Working memory}
+
+The {\sc Oro} server also feature a mechanism to mimic simple forms of
+biological memory.  When new statements are inserted in the knowledge base, a
+\emph{memory profile} is optionally attached to them.
+
+Three such profiles are predefined: {\tt short term}, {\tt episodic} and {\tt
+long term}. They each correspond to a different lifetime for the statements
+(respectively 10 seconds, 5 minutes and no time limit). After this duration,
+the statements are automatically removed from the knowledge base.
+
+This approach is limited. In particular, \emph{episodic} memory primarily
+refers to the semantics of the statements (that is expected to be related to an
+event) and not to a specific life duration.
+
+We rely however on this short term memory for a particular use-case:
+\emph{active concept}. Some modules, like the natural language processor, use
+the {\tt short term} memory profile to mark for a few seconds important
+concepts that are currently manipulated by the robot. For example, if a human
+asks the robot: ``Give me all red objects'', the human, the \concept{Give}
+action, and each red objects that are found are successively marked as
+\emph{active concepts} by inserting statements like \stmt{human type
+ActiveConcept} in the short-term memory (which can be considered, in this case,
+to be a working memory). We use this feature to give a (visual) feedback to the
+users.
 

### 2013-03-22 paper.tex
Section on robot control

@@
 
-\subsection{Modeling the knowledge related to the interaction}
+\subsection{Desires and experiences}
+
+Our robot execution controllers (OpenPRS-based {\sc Shary} or Python-based {\sc
+pyRobots}) have deep integration with the knowledge base. It serves as the primary source of information for semantic-aware decision-making.
 
 We split the interaction situations stemming from the situation assessment and
-communication components in two categories: \emph{desires} and
-\emph{experiences}.
+communication components in two categories: \emph{desires} (performative act)
+and \emph{experiences} (assertive act).
 
-Experiences comprise of emotions and questions
+\emph{Desires} are typically human orders (``Give me that book''). The nature
+of the desired action (to pick, give, look, bring, show...), along with the
+action parametrization (what is acted on? who should perform the action? etc.)
+are extracted from the knowledge base, and either passed to a task planner
+(presented in the next section) or executed if the procedure is directly
+available.
 
-\subsubsection{Event-driven control}
+\emph{Experiences}, on the other hand, comprise of emotions, states and
+questions (when asking a question, we consider the human to be in an
+\emph{interrogative state}). When the knowledge base recognizes that an agent
+\emph{experiences} a particular emotion or state, the execution controller may
+decide to handle it, typically by trying to answer the question or using the
+emotional or physical state as a parameter for subsequent actions. As an example,
+when the speaker says ``I feel tired'', we change the motion planner
+parametrization to lower the effort the human need to provide for the following
+joint manipulation tasks. Note that this example has been implemented as a
+proof-of-concept. We have not yet tried to define a theoretical framework that
+would support action alteration based on the user's experienced states.
+
+\subsection{Event-driven control}
+
+The {\sc Oro} server proposes two paradigms to access its content: RPC-style
+queries (based on the standard SPARQL language) or events. In its simplest
+form, a module can register an event by passing through an event pattern, which
+is basically a partial statement like \stmt{? type Elephant}, and a callback.
+Every time a new instance of elephant appears in the knowledge base, the
+callback is triggered.
+
+This allows us to write reactive robot controllers with a high level of
+expressiveness: for instance, by subscribing to the event \stmt{human1 desires
+?action, ?action type Look, ?action hasGoal myself}, we could trigger a
+behaviour when the human expresses (through dialogue, gestures...) that he
+wants to look at the robot itself.
+
+The robot controller designer does not need to directly care about how this
+\emph{desire} is produced (this is delegated to perception modules), it can
+focuses on the semantic of the desire.
+
+Note also that we take advantage of the reasoning capabilities of the system:
+for example, the goal of the action (\stmt{action hasGoal myself}) may well not
+have been explicitly asserted, but inferred by the reasoner based on other
+assertions.
 
@@
 
-In order to devise how a given goal can be accomplished, the robot has to
-elaborate a plan, \ie a set of actions to be achieved by the robot and its
-human partner(s).  We use in our architecture the HATP planner (for Human Aware
-Task Planner~\cite{Alili2009}).  HATP is based on a Hierarchical Task Network
-(HTN) refinement, which performs an iterative task decomposition into sub-tasks
-until reaching atomic actions.  The \emph{planning domain} defines the set of
-methods describing how to decompose a task and can be seen as the procedural
-knowledge of the robot.
+Our execution controllers rely on symbolic task planning to convert faraway
+desires into a succession of atomic actions. We use in our architecture the
+HATP planner (for Human Aware Task Planner~\cite{Alili2009}).  HATP is based on
+a Hierarchical Task Network (HTN) refinement, which performs an iterative task
+decomposition into sub-tasks until reaching atomic actions.  The \emph{planning
+domain} defines the set of methods describing how to decompose a task and can
+be seen as the procedural knowledge of the robot.
 

### 2013-03-22 paper.tex
Section on LAAS deliberative layer

@@
 
-\cite{Alami2011}
-\ref{fig|archi}
+Fig.~\ref{fig|archi} give an overview of the connections between the
+deliberative components of our architecture~\cite{Alami2011}.
 
@@
 queries previously acquired knowledge to disambiguate and validate the newly
-created symbolic knowledge.
+created symbolic knowledge (see section~\ref{sect|com}).
 
@@
 Regarding the anchoring question, this architecture is bidirectional. The
-components we described provide a \textit{bottom-up} grounding process: SPARK
-and \textsc{Dialogs} constantly build and push new symbolic contents about the
-world to {\sc Oro} where it becomes accessible to decisional layers. In parallel, {\sc Oro}
-relies on reasoning in a \textit{top-down} way to produce new facts that may
-trigger in return physical behaviours.
+components we described provide a \textit{bottom-up} grounding process:
+geometric reasoning and dialogue processing modules constantly build and push
+new symbolic contents about the world to the knowledge base where it becomes
+accessible to decisional layers. In parallel, the knowledge base relies on
+reasoning in a \textit{top-down} way to produce new facts that may trigger in
+return physical behaviours.
 

### 2013-07-08 paper.tex
Minor tuning
Mention the name 'Sally and Ann'

@@
 \usepackage{fancyvrb}
+\usepackage{flushend} % equalize the column length on last page
 
@@
 One classical application of this cognitive skills is the so-called
-\emph{False-Belief} experiment~\cite{Leslie2000}: a child is asked to watch a
-scene where two people, A and B, manipulate objects. Then A leaves and B hides
-away one object. When A comes back, we ask the child ``where do you think A
-will look for the object?''. Before acquiring a theory of mind, children are
-not able to separate their own (true) model of the world (where they know that
-the object was hidden) from the model of A, which contains \emph{false beliefs}
-on the world (A still thinks the object is at its original position since he
-did not see B hiding it). Using seperate knowledge models in the knowledge
-base, we have been able to replicate this experience on our
-robots~\cite{Warnier2012a}.
+\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
+experiment)~\cite{Leslie2000}: a child is asked to watch a scene where two
+people, A and B, manipulate objects. Then A leaves and B hides away one
+object. When A comes back, we ask the child ``where do you think A will
+look for the object?''. Before acquiring a theory of mind, children are not
+able to separate their own (true) model of the world (where they know that
+the object was hidden) from the model of A, which contains \emph{false
+beliefs} on the world (A still thinks the object is at its original
+position since he did not see B hiding it). Using seperate knowledge models
+in the knowledge base, we have been able to replicate this experience on
+our robots~\cite{Warnier2012a}.
 

### 2013-07-08 paper.tex
IROS review: style + grammar

@@
 We have chosen to tackle the challenge from several ends: human-aware
-navigation and motion planning (\cite{Mainprice2011}), situation assessment
-coupled with motion planning (\cite{Mainprice2012}), projection of
-``mightabilities'' that anticipates what surrounding agents may do
-(\cite{Pandey2011}), and the design and deployment of a pervasive
-\emph{knowledge-centric} software architecture.
+navigation and motion planning~\cite{Mainprice2011}, situation assessment
+coupled with motion planning~\cite{Mainprice2012}, projection of
+``mightabilities'' that anticipates what surrounding agents may
+do~\cite{Pandey2011}, and the design and deployment of a pervasive
+\emph{knowledge-oriented} software architecture.
 
-The last item is the focus of this paper: what does ``knowledge-centric''
+The last item is the focus of this paper: what ``knowledge-oriented''
 architecture means, and what are its benefits and drawbacks.
@@
 the functional level. It represents what we call a \emph{knowledge-oriented}
-architecture, built around a central knowledge management tool called the {\sc
-Oro} server.
+architecture, built around a central knowledge management tool called the
+OpenRobot Ontology ({\sc Oro}) server.
 
@@
 used by decision-making components like the robot controller or the task
-planner, and finally, how our knowledge-centric architecture enables the
+planner, and finally, how our knowledge-oriented architecture enables the
 implementation of specific internal cognitive processes.
@@
 This architecture moves away from standard layered approaches. Interactions
-between components are mostly bidirectional and, from the software components
-point of view, we do not introduce layers of abstraction (we do, however, have
-access to the lower level modules of the robot to execute actions, but all
-cognition-related modules reside at the same level). This is especially visible
-for the dialogue input processing. This component does not simply act as an
-alternative perceptual input to the symbolic database, but also actively
-queries previously acquired knowledge to disambiguate and validate the newly
-created symbolic knowledge (see section~\ref{sect|com}).
+between components at the deliberative level are mostly bidirectional and we do
+not introduce layers of abstraction amongst software components\footnote{We do
+have lower-level modules to execute actions or manage sensors, but all
+cognition-related modules reside at the same level.}. This is nicely
+illustrated by the dialogue input processing. This component does not simply
+act as an alternative perceptual input to the symbolic database: it also
+actively queries previously acquired knowledge to disambiguate and validate the
+newly created symbolic knowledge (see section~\ref{sect|com}).
 
@@
 controller that may decide to take an incoming desire into account to create
-its own internal goals. The controller generates and controls intentions from
-these goals with the help of a symbolic task planner, that has also direct
+its own internal goals. The controller generates and manages intentions from
+these goals with the help of a symbolic task planner, that also has direct
 access to the knowledge base.
@@
 server~\cite{Lemaignan2010}) stores knowledge as it is produced by each of the
-deliberative components. It conversely exposes an API to query the knowledge
-base.
+deliberative components. It conversely exposes a {\tt json}-based RPC API to
+query the knowledge base~\cite{lemaignan2012kbapi}.
 
@@
 This knowledge model also largely exclude representation of continuous
-phenomena (like time) or uncertain phenomena. When required, these have to be
-managed by dedicated components.
+phenomena (like time) or uncertain phenomena. When required, these are managed
+inside the corresponding components, and are not exposed at the system level.
 
-Amongst the features of {\sc Oro}, the system can dynamically creates
+Amongst the features of {\sc Oro}, the system can dynamically create
 new independent knowledge models (\ie independent instantiations of the
@@
 their symbolic interpretation. We rely on a dedicated geometric and temporal
-reasoning module (as named on Fig.~\ref{fig|archi}) called SPARK (\emph{SPAtial
-Reasoning \& Knowledge}~\cite{Sisbot2011}). It is a situation assessment
-reasoner that generates symbolic knowledge from the geometry of the
-environment with respect to relations between objects, robots and humans,
-also taking into account the different perspective that each agent has on
-the environment (Fig.~\ref{fig|spark-oro}).
+reasoning module called SPARK (\emph{SPAtial Reasoning \&
+Knowledge}~\cite{Sisbot2011}). It is a situation assessment reasoner that
+generates symbolic knowledge from the geometry of the environment with respect
+to relations between objects, robots and humans (Fig.~\ref{fig|spark-oro}),
+also taking into account the different perspective that each agent has on the
+environment.
 
@@
 symbolic layer. This geometric model is continuously updated at run-time by the
-robot based on its sensors. In our experiments, objects are identified and
-localised through 2D barcodes.  On the other hand, the human pose is tracked
-with a Kinect-like device (assisted by motion capture to accurately track the
-head motion, which is required to compute what the human is looking at).
+robot based on its sensors (in our experiments, objects are identified and
+localised through 2D barcodes, while humans are tracked with Kinect-like
+devices, optionally assisted by motion capture to accurately track the head
+motion, which is required to compute what the human is looking at).
 
@@
 Absolute spatial relations (\concept{isOn}, \concept{isIn}, \concept{isNextTo},
-\concept{isAbove}...) are computed for every pairs of object, taking into
+\concept{isAbove}...) are computed for every pair of objects, taking into
 account relative sizes for relations like \concept{isNextTo}. Other
@@
 computed by placing virtual cameras in the SPARK 3D model, while reachability
-relies on inverse kinematics. Several filters (based on hystersis, expected
+relies on inverse kinematics. Several filters (based on hysteresis, expected
 object behaviour, etc.) allow to partially account for perception
@@
 
-\subsection{Natural language grounding}
+\paragraph*{Natural language grounding}
 
@@
 We detail in~\cite{Lemaignan2011a} the techniques and the tool called
-\texttt{dialogs} that we have developed for English language parsing and
+\texttt{dialogs} that we have developed for natural English language parsing and
 grounding, along with verbalisation and (minimalist) dialogue management.
@@
 structures (for instance, a sentence like ``learn that cats are animals'' is
-converted into \stmt{Cat subClassOf Animal}, interprets common temporal and place adverbs (like \emph{above} or \emph{tomorrow}) and translates to a certain extend \emph{states} (``I'm tired'') into
+converted into \stmt{Cat subClassOf Animal}), interprets common temporal and place adverbs (like \emph{above} or \emph{tomorrow}) and translates to a certain extend \emph{states} (``I'm tired'') into
 \emph{experiences} (\stmt{HUMAN experiences state\_1, state\_1 hasFeature
@@
 
-\subsection{Limits of disambiguation at semantic level}
+\paragraph*{Limits of disambiguation at semantic level}
 
@@
 Note that, while the system benefits from the complementary modalities, they
-are not all required. The dialogue system would for instance run equally well
-with only the verbal modality.
+are not all required. The dialogue system would for instance happily run with
+only the verbal modality, at the cost of weaker interaction.
 
@@
 queries (based on the standard SPARQL language) or events. In its simplest
-form, a module can register an event by passing through an event pattern, which
-is basically a partial statement like \stmt{? type Elephant}, and a callback.
-Every time a new instance of elephant appears in the knowledge base, the
-callback is triggered.
+form, a module can subscribe to an event by passing through an event pattern
+(in its basic form, a partial statement like \stmt{? type Elephant}) and a
+callback.  Every time a new instance of elephant appears in the knowledge base,
+the callback is triggered.
 
@@
 \emph{desire} is produced (this is delegated to perception modules), it can
-focuses on the semantic of the desire.
+focus on the semantic of the desire.
 
 Note also that we take advantage of the reasoning capabilities of the system:
-for example, the goal of the action (\stmt{action hasGoal myself}) may well not
-have been explicitly asserted, but inferred by the reasoner based on other
-assertions.
+for example, the goal of the action (\stmt{action hasGoal myself}) may not be
+explicitly asserted, but inferred by the reasoner based on other assertions.
 
@@
 and look for correlations with the cognitive load of a human partner (measured
-via EEG of qualitative feedback).
+via EEG or qualitative feedback).
 
@@
 
-The {\sc Oro} server also feature a mechanism to mimic simple forms of
+The {\sc Oro} server also feature a mechanism to mimic minimalistic forms of
 biological memory.  When new statements are inserted in the knowledge base, a
@@
 We rely however on this short term memory for a particular use-case:
-\emph{active concept}. Some modules, like the natural language processor, use
+\emph{active concepts}. Some modules, like the natural language processor, use
 the {\tt short term} memory profile to mark for a few seconds important
@@
 ActiveConcept} in the short-term memory (which can be considered, in this case,
-to be a working memory). We use this feature to give a (visual) feedback to the
-users.
+to be a working memory). We use this feature to trace certain knowledge-related
+processes.
 
@@
 
-We conclude this article with three different perspectives on the knowledge-centric
+We conclude this article with three different perspectives on the knowledge-oriented
 techniques we have presented here: first, the \emph{architect perspective}:
-what are the pro and cons of our approach from an engineer point of view. Then,
+what are the pro and cons of our approach from an engineer's point of view. Then,
 the \emph{logician perspective}: does the promise of automated reasoning for
@@
 has one important consequence from the perspective of the system architecture:
-modules interfaces are only defined in term of knowledge consumption and/or
+interfaces between modules are only defined in term of knowledge consumption and/or
 production. This brings good decoupling properties. For instance, the output of
@@
 produces or needs which knowledge, instead of which module produces which
-service. We found this approach to be very helpful when implementing cognitive
+service. We found this approach to be helpful when implementing cognitive
 functions like the theory of mind, the perspective taking or the natural
@@
 modules (which is even easier in our case since we rely on a central knowledge
-management), one can trace the interactions between the robot systems at a high
+management), one can trace the interactions between the robot subsystems at a high
 abstraction level. We will come back below on the significance of this property
@@
 
-The knowledge-centric architecture also has downsides. We can name here a few.
+The knowledge-oriented architecture also has downsides. We can name here a few.
 
@@
 
-On the sensing side also we have in certain cases to bypass the semantic layer
+Also on the sensing side, we have in certain cases to bypass the semantic layer
 (extension of the symbolic task planner with geometric constraints is one
@@
 
-And obviously, many aspects like uncertainty management, modal logics,
-non-monotonic reasoning,... are not well addressed with the Description Logic
+Many aspects like uncertainty management, modal logics,
+non-monotonic reasoning are not well addressed with the Description Logic
 formalism, and remain to be explored to gain the level of expressiveness
-required to cover many of the more complex human-robot interaction scenarii.
+required to cover many of the more complex human-robot interaction scenarios.
 
@@
 simple yet non-trivial reasoning. As a matter of fact, only very few of such
-prototypical reasoning cases where positively identified in our scenarii and experiments
+prototypical reasoning cases were positively identified in our scenarios and experiments
 (and consequently implemented as rules in {\sc Oro}).
@@
 The experiments we have conducted are also likely to have too simplistic
-semantics to let complex reasoning needs to emerge. Scenarii with more complex
+semantics to let complex reasoning needs to emerge. Scenarios with more complex
 semantics would be desirable to better stress the expressiveness and inference
@@
 constraints, transitivity, etc.) and encode a large amount of common-sense
-knowledge that can not be otherwise asserted. These trivial inferences are all
-the more important that an expressive knowledge representation language is
-used: when a language like OWL allows to directly represent high-level concepts
-like partitions, cardinality restrictions, properties' ranges and domains, it
-leads to a more implicit (because more abstract) description of the vocabulary
-that in turn requires more underlying reasoning. With the progress in the
+knowledge that can not be otherwise conveniently asserted. These trivial
+inferences are all the more important that an expressive knowledge
+representation language is used. Since a language like OWL allows it to
+directly represent high-level concepts like partitions, cardinality
+restrictions, properties' ranges and domains, it leads to a more
+\emph{implicit} description of the available knowledge.  This in turn pushes
+for a lot of ``trivial'' reasoning to make the knowledge pool explicit, and
+hence largely reusable by the deliberative subsystems. With the progress in the
 understanding of the relations between expressiveness and (tractable)
 satisfiability, along with the progress of reasoners, more and more of the
-inferences do not need to be explicit anymore, and consequently are delegated
+assertions do not need to be explicit anymore, and consequently are delegated
 to the reasoner.

### 2013-07-08 paper.tex
Remove the section on the blind and deaf robot

@@
 not only to enable rich human-robot interaction, but also as a broader approach
-to information alignment and fusion in complex robotic systems.  The
-versatility of this paradigm could be illustrated by a simple imaginary
-scenario with a blind robot and a deaf robot. The blind robot does not see (no
-cameras or alike), but someone can verbally describe a scene to it. On the
-other hand, the deaf robot has a good vision system, but cannot process verbal
-input.  Without any changes to the software architecture that we described,
-control modules of both robots could equally perform the same tasks since all
-the knowledge is abstracted and centralised (note that to actually implement
-this imaginary situation, the blind robot would of course need \textit{a
-priori} 3D models of objects talked about to enable planning or pick and place
-actions, and the deaf robot would require at least some gesture interpretation
-to understand orders).
+to information alignment and fusion in complex robotic systems.
 

### 2013-07-08 paper.tex
Summarized SPARK section

@@
 
-\subsection{Computed symbolic relations}
+\paragraph*{Computed symbolic relations} Over twenty physical relations are
+continuously computed by SPARK: spatial relations, both agent-independent (like
+\concept{isOn}, \concept{isNextTo} or \concept{isMoving}), and agent-dependent
+(like \concept{isNear}, \concept{leftTo}), and also affordances:
+\concept{sees}, \concept{looksAt}, \concept{pointsAt}, \concept{canReach}.
+Visibility (with two levels corresponding to the field of view and the field of
+attention) and pointing are computed by placing virtual cameras in the SPARK 3D
+model, while reachability relies on inverse kinematics. Several filters (based
+on hysteresis, expected object behaviour, etc.) allow to partially account for
+perception noise~\cite{Warnier2012a}.
 
-\subsubsection{Agent-independent locations}
-
-Absolute spatial relations (\concept{isOn}, \concept{isIn}, \concept{isNextTo},
-\concept{isAbove}...) are computed for every pair of objects, taking into
-account relative sizes for relations like \concept{isNextTo}. Other
-agent-independent properties like \concept{isMoving} are also computed.
-
-\subsubsection{Agent-dependent placements}
-
-The space surrounding each agent is divided into 16 regions (like
-\emph{back-left}, \emph{front-near}, etc.). Based on this partition, symbolic
-relations between each pair (agent, object) is computed and added to the
-knowledge base.
-
-\subsubsection{Affordances}
-
-SPARK also computes affordances: \concept{sees}, \concept{looksAt},
-\concept{pointsAt}, \concept{canReach}. Visibility (with two levels
-corresponding to the field of view and the field of attention) and pointing are
-computed by placing virtual cameras in the SPARK 3D model, while reachability
-relies on inverse kinematics. Several filters (based on hysteresis, expected
-object behaviour, etc.) allow to partially account for perception
-noise~\cite{Warnier2012a}.
-
-\subsection{Ego-centric and allo-centric frames}
-
-SPARK enables \emph{perspective-taking}: spatial relations between entities can
-be computed from different viewpoints, which let the robot build for each
-agents it interacts with a different, \emph{perspective-aware} symbolic model
-of the environment. These models are separately stored in the knowledge base.
+\paragraph*{Ego-centric and allo-centric frames} SPARK enables
+\emph{perspective-taking}: spatial relations between entities can be computed
+from different viewpoints, which let the robot build for each agents it
+interacts with a different, \emph{perspective-aware} symbolic model of the
+environment. These models are separately stored in the knowledge base.
 
@@
 As a result, the robot stores models on the environment either in the
-\emph{ego-centric} reference frame (from the robot perspective) or in
-addressee-centred (\ie, \emph{allo-centric}) frames.
+\emph{ego-centric} reference frame (from the robot perspective) or in the
+\emph{allo-centric} frame (addressee-centred).
 

### 2013-07-08 paper.tex
Added a tentative conclusion

@@
 
+\section{Conclusion}
+
+This article does not claim to provide an exhaustive assessment of the
+strengths and weaknesses of explicit knowledge management in a large robotic
+architecture. Most of the ``lessons'' that we have ``learned'' are not new, and
+the various scientific communities (task planning, architecture control, NLP,
+geometric reasoning, social psychology...) that are related to human-robot
+interaction and robotic cognition already know them.
+
+One of the difficult challenge of cognitive robotics, however, is to overcome
+the relative isolation of each of these communities, to build an embodied
+agent, able to autonomously interact.
+
+In this respect, our main contribution is to propose a critical look at a
+viable robotic architecture that covers many of the facets of cognitive
+robotics and that is actually deployed on several robots.
+
+And if we had a single ``lesson'' to retain, it would likely be that one:
+shifting from a robotic system design based on \emph{modules and APIs} to a
+system based on \emph{modules and human-level semantic interfaces} has been the
+main facilitator both for glueing together deliberative heterogeneous modules
+in a meaningful way, and quickly developing new upper-cognition abilities,
+including natural dialogue and theory of mind.
+
 \section*{Acknowledgment}

### 2013-07-09 paper.tex
Make the scope of the article clear
(knowledge for integration, no literature review, no novelty at component levels)

@@
 
-\subsection{Deployment on robots}
+\subsection{Scope of the article}
+
+In essence, this article portrays our use of \emph{explicit semantic
+interfaces} to integrate several deliberative components into a decisional
+architecture for our robots. It reports on our experience with acquiring,
+managing and reusing grounded knowledge in the context of human-robot
+interaction.
+
+Unlike previous publications by the authors~\cite{Pandey2011, Lemaignan2010,
+Ros2010b, Lemaignan2011a, Warnier2012a, Alami2011, Sisbot2011, Alili2009} that
+introduced the deliberative modules of our robots independently from each
+others, this contribution is an account of the importance of explicit knowledge
+manipulation when \emph{integrating} them into a large cognitive architecture.
 
@@
 reported experiments in~\cite{Lemaignan2010, Ros2010b, Lemaignan2011a,
-Warnier2012a}). Also, videos presenting some of these experimental results can
+Warnier2012a}). Videos presenting some of these experimental results can
 be watched online at~\url{http://www.laas.fr/~slemaign}.
+
+Also note that this paper does not provide extensive review of existing
+literature. We kindly refer the reader to the aforementioned articles for
+in-depth discussion of related works in the different fields of cognitive
+architectures for robots.
 

### 2013-07-09 paper.tex
IROS review: clarify relation with BDI

@@
 
-Our architecture relates but is to be distinguished from \emph{Beliefs,
-Desires, Intentions} (BDI) architectures. BDI architectures are primarily
-focused on \emph{practical reasoning}, \ie the process of deciding, step by
-step, which action to perform to reach a goal (as summarised by
-Woolridge~\cite{Woolridge1999}). The management of the interaction between
-knowledge (the beliefs) and task and plan representation and execution (the
-desires and the intentions) is central, and aims at selecting at each step the
-best subgoal. It becomes then an intention that the robot commits to.
+Our architecture relates to \emph{Beliefs, Desires, Intentions} (BDI)
+architectures. BDI architectures are primarily focused on \emph{practical
+reasoning}, \ie the process of deciding, step by step, which action to perform
+to reach a goal (as summarised by Woolridge~\cite{Woolridge1999}). The
+management of the interaction between knowledge (the beliefs) and task and plan
+representation and execution (the desires and the intentions) is central, and
+aims at selecting at each step the best subgoal. It becomes then an intention
+that the robot commits to.
 
 This interaction between knowledge and actions is also central to our approach
-(as for any cognitive system), but task representation and task execution is
-not seen as a monolithic, central function: it is one of the activities of the
-robot, actually split between communication components (that can acquire
+(as for any cognitive system): it is one of the activities of the
+robot, actually shared between communication components (that can acquire
 desires from interaction with agents, amongst other things) and an execution
@@
 
-The architecture is not only focused on this workload, and other
-activities are conducted without being explicitly considered as desires:
-assessment of the situation and the environment, dialogue (including
+Our architecture is however \emph{not built around} this activity only. Other
+activities are conducted in parallel, without being explicitly considered as
+desires: assessment of the situation and the environment, dialogue (including
 performative dialogue that can possibly change the internal state of the robot,
-but does not lead to the creation of desires,  like question answering or
+but does not lead to the creation of desires,like question answering or
 statement assertion), various background monitoring and recognition tasks, etc.
@@
 
-\subsection{Knowledge model}
+\subsection*{Knowledge model}
 

### 2013-07-09 paper.tex
IROS review: extend discussion on knowledge model

@@
 Relying on RDF triples and Description Logics has major advantages like the
-availability of numerous open-source libraries to manipulate the ontology,
+availability of numerous mature open-source libraries to manipulate the ontology,
 interoperability with several major on-line knowledge bases (like {\sc
@@
 
-It has also notable limitations: RDF triples imply only binary predicates
+It has also notable limitations, both fundamental (the suitability
+of Description Logics when reasoning on --typically non-monotonic-- commonsense
+knowledge is questionable) and practical: RDF triples imply only binary predicates
 (\stmt{subject predicate object}), which constrains the expressiveness of the
@@
 This knowledge model also largely exclude representation of continuous
-phenomena (like time) or uncertain phenomena. When required, these are managed
-inside the corresponding components, and are not exposed at the system level.
+phenomena (like time) or uncertain phenomena. When required (for instance for
+action recognition), these are managed inside the corresponding components, and
+are not exposed at the semantic level.
 
-Amongst the features of {\sc Oro}, the system can dynamically create
-new independent knowledge models (\ie independent instantiations of the
-ontology). This feature is used for instance to store separately the beliefs of
-each of the agents the robot interacts with (see section~\ref{sect|tom}).
+Tools to manipulate and reason over ontologies are readily available, mature,
+and already well accepted in the robotic community~\cite{Tenorth2009a,
+Lim2011}. While alternatives like \emph{Answer Set Programming} have also been
+successfully investigated in robotics~\cite{Chen2010,Erdem2012}, in particular
+to deal with non-monotonic reasoning, we did not actually hit any brick wall
+while working with OWL ontologies. We may reconsider this choice at a later
+stage, but until now it has proven an effective framework to quickly explore
+implementations of new cognitive abilities (for instance, it has been
+conceptually and technically easy to add support for independent knowledge
+models, one per the agents the robot interacts with --see
+section~\ref{sect|tom}).
+
+Besides, because ontologies and RDF statements are relatively simple concepts
+to grasp, it also effectively helped to grow awareness amongst colleagues on
+the significance of the ``semantic level'' when developing new components for
+the robot.
 

### 2013-07-09 paper.tex
IROS review: clarify/refocused conclusion

@@
 
-\subsection{Knowledge-oriented architecture}
-
 Altogether, the components we have presented compose an architecture that we
@@
 
-    \item{The first two points enable both a loosely-coupled architecture where
+    \item{The first two points enable a loosely-coupled architecture where
     modules can very easily be removed or replaced by other ones as long as
@@
 
-    \item{and a \emph{symbolic}, reactive, event-driven approach to robot control.
+    \item{We adopt a \emph{symbolic}, reactive, event-driven approach to robot control.
     By managing events at the same level as the reasoner, we take full
@@
 
-We believe that this \emph{knowledge-oriented} approach has a strong potential
-not only to enable rich human-robot interaction, but also as a broader approach
-to information alignment and fusion in complex robotic systems.
+Those items are not new \emph{per se}. It seems however interesting to
+underline the shift of focus this implies during the design and integration
+phases. We can adopt three perspectives to discuss the techniques we have
+presented here: first, the \emph{architect perspective}: what are the pro and
+cons of our approach from an engineer's point of view. Then, the \emph{logician
+perspective}: does the promise of automated reasoning for robots really comes
+to fruition with ontologies? And finally the \emph{cognitician perspective} or
+how explicit knowledge manipulation endows our robots with new cognitive
+capabilities.
 
-\subsection{Three perspectives on knowledge manipulation for robots}
-
-We conclude this article with three different perspectives on the knowledge-oriented
-techniques we have presented here: first, the \emph{architect perspective}:
-what are the pro and cons of our approach from an engineer's point of view. Then,
-the \emph{logician perspective}: does the promise of automated reasoning for
-robots really comes to fruition with ontologies? And finally the
-\emph{cognitician perspective} or how explicit knowledge manipulation endows
-our robots with new cognitive capabilities.
-
-\subsubsection{The Architect view: loose coupling and modalities merging}
+\subsection{The Architect view: loose coupling and modalities merging}
 
@@
 
-\subsubsection{The Logician view: the importance of trivial inferences}
+\subsection{The Logician view: the importance of trivial inferences}
 
@@
 
-\subsubsection{The Cognitician view: palpable knowledge and semantic thinking}
+\subsection{The Cognitician view: palpable knowledge and semantic thinking}
 
@@
 One of the difficult challenge of cognitive robotics, however, is to overcome
-the relative isolation of each of these communities, to build an embodied
-agent, able to autonomously interact.
+the relative isolation of each of these communities to build an autonomous
+embodied agent, able to interact with humans. This contribution modestly offers
+a critical look at one instance of such a ``real-world'' robotic architecture,
+that tackles several of the facets of cognitive robotics.
 
-In this respect, our main contribution is to propose a critical look at a
-viable robotic architecture that covers many of the facets of cognitive
-robotics and that is actually deployed on several robots.
-
-And if we had a single ``lesson'' to retain, it would likely be that one:
-shifting from a robotic system design based on \emph{modules and APIs} to a
-system based on \emph{modules and human-level semantic interfaces} has been the
-main facilitator both for glueing together deliberative heterogeneous modules
-in a meaningful way, and quickly developing new upper-cognition abilities,
-including natural dialogue and theory of mind.
+If we had a single ``lesson'' to retain, it would likely be that one: shifting
+from a robotic design based on \emph{modules and APIs} to a system based on
+\emph{modules and human-level semantic interfaces} has been the main
+facilitator both for glueing together heterogeneous deliberative modules in a
+\emph{meaningful} way, and quickly developing prototypes for new
+upper-cognition abilities, like natural dialogue and theory of mind.
 

### 2013-07-10 paper.tex
[IROS review] various minor improvements

@@
 
-The system supports quantification (``give me \{a | the | some | all | any | ...\} can''),
-thematic roles (action-specific predicates that qualify the actions), interactive disambiguation (the robot asks
-questions when it needs more information), anaphora resolution (``give \emph{it} to me'') based on
+The system supports quantification (``give me \{a | the | some | all | any |
+...\} can''), thematic roles (action-specific predicates that qualify the
+actions), interactive disambiguation (the robot asks questions when it needs
+more information), anaphora resolution (``give \emph{it} to me'') based on
 dialogue history. It also permits knowledge extension by learning new semantic
 structures (for instance, a sentence like ``learn that cats are animals'' is
-converted into \stmt{Cat subClassOf Animal}), interprets common temporal and place adverbs (like \emph{above} or \emph{tomorrow}) and translates to a certain extend \emph{states} (``I'm tired'') into
-\emph{experiences} (\stmt{HUMAN experiences state\_1, state\_1 hasFeature
-tired}).
+converted into \stmt{Cat subClassOf Animal}), interprets common temporal and
+place adverbs (like \emph{above} or \emph{tomorrow}) and translates to a
+certain extend \emph{states} (``I'm tired'') into \emph{experiences}
+(\stmt{HUMAN experiences state\_1, state\_1 hasFeature tired}).
 
@@
 Our robot execution controllers (OpenPRS-based {\sc Shary} or Python-based {\sc
-pyRobots}) have deep integration with the knowledge base. It serves as the primary source of information for semantic-aware decision-making.
+pyRobots}) have deep integration with the knowledge base. It serves as the
+primary source of information for semantic-aware decision-making.
 
@@
 decomposition into sub-tasks until reaching atomic actions. The \emph{planning
-domain} defines the set of methods describing how to decompose a task and can
-be seen as the procedural knowledge of the robot.
+domain} defines the set of methods describing how to decompose a task and
+represents the procedural knowledge of the robot.
 
@@
 tasks pre- and postconditions) is stored in a specific format, outside of the
-knowledge repository, and the plans are directly communicated to the robot
-controller. From our experience, the OWL formalism is not a satisfying
-representation language for concepts that are both tightly bound to time
-representation and non-monotonic.
+central declarative knowledge repository, and the plans are directly
+communicated to the robot controller. Thus, like many other cognitive
+architectures, we have independent declarative and procedural knowledge stores.
 
@@
 \label{sect|intern}
-
-\subsection{Cognitive load}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=\columnwidth]{figs/cognitive_load.png}
-        \caption{This diagram shows the average number of interactions with the
-        knowledge base during a one-hour long experiment with the robot.
-        Interactions are either knowledge alteration (addition/removal of
-        statements) or queries. They are averaged on a sliding window of 60 seconds.}
-        \label{fig|cognitiveload}
-\end{figure}
-
-Fig.~\ref{fig|cognitiveload} is a plot of the log of interactions between the
-robot modules and the knowledge base during a one-hour long experiment. Because
-most of the communication between modules goes in fact through the knowledge
-base in our architecture (Fig.~\ref{fig|archi}), it actually pictures the
-intensity of internal information exchanges between the different deliberative
-component of the robot over time. Since communication actually relies on
-grounded semantic information, we can call this level of interaction the
-\emph{cognitive load} of the robot.
-
-This simplistic measurement needs to be refined to take into account the fact
-that the raw number of RPC calls to the knowledge base only imperfectly
-reflects the actual cognitive processes running on, but it opens interesting
-perspective nevertheless. It would be interesting for example to compare
-and look for correlations with the cognitive load of a human partner (measured
-via EEG or qualitative feedback).
 
@@
 This skill, which is essential to understand others' perspectives during
-interactions, appears in children around their third year. It supposes the ability to build, store and retrieve separate models of the knowledge of the interactors.
+interactions, appears in children around their third year. It supposes the
+ability to build, store and retrieve separate models of the knowledge of the
+interactors.
 
@@
 processes.
+
+\subsection{Cognitive activity}
+
+\begin{figure}
+        \centering
+        \includegraphics[width=\columnwidth]{figs/cognitive_load.png}
+        \caption{This diagram shows the average number of interactions with the
+        knowledge base during a one-hour long experiment with the robot.
+        Interactions are either knowledge alteration (addition/removal of
+        statements) or queries. They are averaged on a sliding window of 60 seconds.}
+        \label{fig|cognitiveload}
+\end{figure}
+
+Fig.~\ref{fig|cognitiveload} is a plot of the interactions (knowledge
+alterations or knowledge queries) between the robot's modules and the central
+knowledge base during a one-hour long experiment. Because most of the
+communication between modules are in fact supported by the knowledge base in
+our architecture (Fig.~\ref{fig|archi}), it reflects well the exchanges of
+semantic informations within the deliberative layer of the robot, which we can
+tentatively call the \emph{cognitive activity} of the robot.
+
+This simplistic measurement needs to be refined to take into account the fact
+that the raw number of RPC calls to the knowledge base only imperfectly
+reflects the actual cognitive processes running on, but it opens interesting
+perspectives nevertheless. We could for instance investigate possible
+correlations with the cognitive activity of a human partner (measured via EEG
+or qualitative feedback).
 

### 2013-07-11 paper.tex
Julia's review
Including minor update to figures

@@
 prerequisites of these challenges, and giving them experimental reality on the
-robots: what is required for sentences like ``Let's dress the table together''
+robots: what is required for sentences like ``Let's set the table together''
 to be understood by the robot, correctly interpreted in the spatial and
@@
 The last item is the focus of this paper: what ``knowledge-oriented''
-architecture means, and what are its benefits and drawbacks.
+architecture means, and what its benefits and drawbacks are.
 
@@
 
-Unlike previous publications by the authors~\cite{Pandey2011, Lemaignan2010,
-Ros2010b, Lemaignan2011a, Warnier2012a, Alami2011, Sisbot2011, Alili2009} that
-introduced the deliberative modules of our robots independently from each
-others, this contribution is an account of the importance of explicit knowledge
-manipulation when \emph{integrating} them into a large cognitive architecture.
+Unlike previous publications by the authors that introduced the deliberative
+modules of our robots independently from each other, this contribution is an
+account of the importance of explicit knowledge manipulation when
+\emph{integrating} them into a large cognitive architecture.
 
@@
 reported experiments in~\cite{Lemaignan2010, Ros2010b, Lemaignan2011a,
-Warnier2012a}). Videos presenting some of these experimental results can
-be watched online at~\url{http://www.laas.fr/~slemaign}.
+Warnier2012a, Sisbot2011}). Videos presenting some of these experimental
+results can be watched online at~\url{http://www.laas.fr/~slemaign}.
 
@@
 
-The following sections go over each of the facets of this deliberative layer:
+The following sections present each of the facets of this deliberative layer:
 how knowledge is produced asynchronously from geometric reasoning, how the
@@
 such a knowledge-oriented architecture, from three perspectives: the software
-architect, the IA expert and the cognitician.
+architect, the AI expert and the cognitician.
 
@@
 
-Fig.~\ref{fig|archi} give an overview of the connections between the
+Fig.~\ref{fig|archi} gives an overview of the connections between the
 deliberative components of our architecture~\cite{Alami2011}.
@@
 illustrated by the dialogue input processing. This component does not simply
-act as an alternative perceptual input to the symbolic database: it also
+act as an alternative perceptual input to the symbolic database; it also
 actively queries previously acquired knowledge to disambiguate and validate the
@@
 reasoning}, \ie the process of deciding, step by step, which action to perform
-to reach a goal (as summarised by Woolridge~\cite{Woolridge1999}). The
+to reach a goal (as summarized by Woolridge~\cite{Woolridge1999}). The
 management of the interaction between knowledge (the beliefs) and task and plan
@@
 
-Our architecture is however \emph{not built around} this activity only. Other
+This activity is however not the backbone of our architecture. Other
 activities are conducted in parallel, without being explicitly considered as
@@
 performative dialogue that can possibly change the internal state of the robot,
-but does not lead to the creation of desires,like question answering or
+but does not lead to the creation of desires, like question answering or
 statement assertion), various background monitoring and recognition tasks, etc.
@@
 accessible to decisional layers. In parallel, the knowledge base relies on
-reasoning in a \textit{top-down} way to produce new facts that may trigger in
-return physical behaviours.
+reasoning in a \textit{top-down} way to produce new facts that may in return
+trigger physical behaviours.
 
@@
 In our architecture (Fig.~\ref{fig|archi}), knowledge manipulation relies on a
-\emph{semantic blackboard}: one central server (the {\sc Oro}
+\emph{semantic blackboard}: a central server (the {\sc Oro}
 server~\cite{Lemaignan2010}) stores knowledge as it is produced by each of the
@@
 
-Relying on RDF triples and Description Logics has major advantages like the
+Relying on RDF triples and Description Logics has advantages such as the
 availability of numerous mature open-source libraries to manipulate the ontology,
@@
 
-It has also notable limitations, both fundamental (the suitability
+It also has notable limitations, both fundamental (the suitability
 of Description Logics when reasoning on --typically non-monotonic-- commonsense
@@
 (\stmt{subject predicate object}), which constrains the expressiveness of the
-system or leads to cumbersome reifications. Alternative exists (like {\sc
-KnowRob}~\cite{Tenorth2009a}) that mixes RDF with more expressive logic
+system or leads to cumbersome reifications. Alternatives exist (like {\sc
+KnowRob}~\cite{Tenorth2009a}) that mix RDF with more expressive logic
 languages like {\sc Prolog}, at the price, however, of other limitations, like
@@
 standard experiment (about 100 classes and 200 instances), classification
-typically take about 100ms, which becomes problematic during interactions.
+typically takes about 100ms, which becomes problematic during interactions.
 Besides, the performances are difficult to predict, since a seemingly
@@
 
-This knowledge model also largely exclude representation of continuous
+This knowledge model also largely excludes representation of continuous
 phenomena (like time) or uncertain phenomena. When required (for instance for
@@
         \centering
-        \includegraphics[width=\columnwidth]{spark-oro}
+        \includegraphics[width=\columnwidth]{spark-oro.png}
     \caption{Functional overview of knowledge base ({\sc Oro} server, top part) and the geometric situation assessment module (\emph{SPARK}, bottom part)}
@@
 \emph{perspective-taking}: spatial relations between entities can be computed
-from different viewpoints, which let the robot build for each agents it
-interacts with a different, \emph{perspective-aware} symbolic model of the
-environment. These models are separately stored in the knowledge base.
+from different viewpoints, which let the robot build a different,
+\emph{perspective-aware} symbolic model of the environment for each agent it
+interacts with. These models are separately stored in the knowledge base.
 
-This allows us to deal with ambiguities that arises when one speaker refers to
+This allows us to deal with ambiguities that arise when one speaker refers to
 an object within a reference system (or changes the reference system, \ie
@@
 
-As a result, the robot stores models on the environment either in the
+As a result, the robot stores models of the environment either in the
 \emph{ego-centric} reference frame (from the robot perspective) or in the
@@
 Natural language processing is one of the fields of human-robot interaction
-where the introduction of the semantic layer has been most beneficial.
-We detail in~\cite{Lemaignan2011a} the techniques and the tool called
+for which the introduction of the semantic layer has been most beneficial.
+In~\cite{Lemaignan2011a} we detail the techniques and the tool called
 \texttt{dialogs} that we have developed for natural English language parsing and
@@
 with Google speech recognition) is parsed into a grammatical structure, and
-each sentence atoms are resolved with the help of the ontology to ground
+atoms of each sentence are resolved with the help of the ontology to ground
 concepts like objects (\ie when a user says ``pick the can'', resolve to which
@@
 
-Besides, even knowledge that belongs the \emph{common knowledge} may not be
+Besides, even knowledge that belongs to the \emph{common knowledge} may not be
 appropriate in a given interaction context. For instance, the system may
@@
 the different communication modalities (\emph{explicit} like verbal, deictic or
-based on gaze, or \emph{implicit} like postures) are presented in an
-homogeneous way.
-
-The dialogue grounding process makes use of them at two distinct levels.
+based on gaze, or \emph{implicit} like postures) are presented in a homogeneous
+way. The dialogue grounding process makes use of them at two distinct levels.
 
@@
 presence and value of specific facts: for instance, when a set of instances
-match a category (the human says ``give me the bottle'' and the robot knows
+matches a category (the human says ``give me the bottle'' and the robot knows
 about three bottles), the module may decide (it actually depends on the
 quantifier preceding the class) to discard some of them based on their
-\emph{visibility} by the speaker (implicit communication context built on the
+\emph{visibility} for the speaker (implicit communication context built on the
 human posture).
@@
 discrimination possibilities are available: for instance, if reachability is
-available, the robot may asks ``do you mean the bottle that is accessible to
+available, the robot may ask ``do you mean the bottle that is accessible to
 me?'' to discriminate between the three bottles. That way, procedures relying
@@
 when the speaker says ``I feel tired'', we change the motion planner
-parametrization to lower the effort the human need to provide for the following
+parametrization to lower the effort the human needs to provide for the following
 joint manipulation tasks. Note that this example has been implemented as a
@@
 (in its basic form, a partial statement like \stmt{? type Elephant}) and a
-callback.  Every time a new instance of elephant appears in the knowledge base,
+callback.  Each time a new instance of elephant appears in the knowledge base,
 the callback is triggered.
@@
 The robot controller designer does not need to directly care about how this
-\emph{desire} is produced (this is delegated to perception modules), it can
+\emph{desire} is produced (this is delegated to perception modules), he can
 focus on the semantic of the desire.
@@
 Second remark, after some research (see appendix B of~\cite{Lemaignan2012a} for
-a detailed discussion)  we have decided not to represent neither the planning
+a detailed discussion)  we have decided to represent neither the planning
 domain nor the resulting plans in the knowledge base: the planning domain (with
-tasks pre- and postconditions) is stored in a specific format, outside of the
+task pre- and postconditions) is stored in a specific format, outside of the
 central declarative knowledge repository, and the plans are directly
@@
 ability that a subject possesses to represent the mental state of another
-agent, possibly including knowledge that contradicts the subject own model: for
-example, a book can be at the same time \emph{visible} by myself, and \emph{not
-visible} by you.
+agent, possibly including knowledge that contradicts the subject's own model: for
+example, a book can be at the same time \emph{visible} for myself, and \emph{not
+visible} for you.
 
-This skill, which is essential to understand others' perspectives during
-interactions, appears in children around their third year. It supposes the
+Children develop this skill, which is essential to understand others' perspectives during
+interactions, around the age of three. It supposes the
 ability to build, store and retrieve separate models of the knowledge of the
@@
 
-One classical application of this cognitive skills is the so-called
+One classical application of this cognitive skill is the so-called
 \emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
@@
 beliefs} on the world (A still thinks the object is at its original
-position since he did not see B hiding it). Using seperate knowledge models
-in the knowledge base, we have been able to replicate this experience on
+position since he did not see B hiding it). Using separate knowledge models
+in the knowledge base, we have been able to replicate this experience with
 our robots~\cite{Warnier2012a}.
@@
 
-The {\sc Oro} server also feature a mechanism to mimic minimalistic forms of
+The {\sc Oro} server also features a mechanism to mimic minimalistic forms of
 biological memory.  When new statements are inserted in the knowledge base, a
@@
 action, and each red objects that are found are successively marked as
-\emph{active concepts} by inserting statements like \stmt{human type
+\emph{active concepts} by inserting statements such as \stmt{human type
 ActiveConcept} in the short-term memory (which can be considered, in this case,
@@
 knowledge base during a one-hour long experiment. Because most of the
-communication between modules are in fact supported by the knowledge base in
+communication between the modules are in fact supported by the knowledge base in
 our architecture (Fig.~\ref{fig|archi}), it reflects well the exchanges of
@@
 that the raw number of RPC calls to the knowledge base only imperfectly
-reflects the actual cognitive processes running on, but it opens interesting
+reflect the actual cognitive processes running on, but it opens interesting
 perspectives nevertheless. We could for instance investigate possible
@@
     \item{Knowledge is explicitly stored in one central and consistent
-    repository of facts, accessible by all modules.}
+    repository of facts, accessible for all modules.}
 
@@
     \item{The first two points enable a loosely-coupled architecture where
-    modules can very easily be removed or replaced by other ones as long as
+    modules can be removed or replaced easily by other ones as long as
     they share the same semantics (modules are defined by the knowledge they
@@
 phases. We can adopt three perspectives to discuss the techniques we have
-presented here: first, the \emph{architect perspective}: what are the pro and
+presented here: first, the \emph{architect perspective}: what are the pros and
 cons of our approach from an engineer's point of view. Then, the \emph{logician
-perspective}: does the promise of automated reasoning for robots really comes
+perspective}: does the promises of automated reasoning for robots really come
 to fruition with ontologies? And finally the \emph{cognitician perspective} or
@@
 has one important consequence from the perspective of the system architecture:
-interfaces between modules are only defined in term of knowledge consumption and/or
+interfaces between modules are only defined in terms of knowledge consumption and/or
 production. This brings good decoupling properties. For instance, the output of
@@
 the same chunks of knowledge. Only a thin protocol layer remains to get the
-modules to interact together. This allows us to implement (otherwise difficult)
+modules to interact with each other. This allows us to implement (otherwise difficult)
 merging of radically different perception modalities like vision (the robot
@@
 de facto} defined by the semantics of the information they
-convey. This allows the system designers to think in term of which module
+convey. This allows the system designers to think in terms of which module
 produces or needs which knowledge, instead of which module produces which
@@
 
-Lastly, another good property of knowledge-based communication between module
+Finally, another good property of knowledge-based communication between modules
 is the \emph{cognitive observability} of the system: by logging queries between
@@
 management), one can trace the interactions between the robot subsystems at a high
-abstraction level. We will come back below on the significance of this property
-from the cognitive point of view.
+abstraction level. We address the significance of this property
+from the cognitive point of view below.
 
-The knowledge-oriented architecture also has downsides. We can name here a few.
+The knowledge-oriented architecture also has shortcomings. We can name here a few.
 
@@
 
-Also on the sensing side, we have in certain cases to bypass the semantic layer
+Also on the sensing side, we have to bypass the semantic layer in certain cases
 (extension of the symbolic task planner with geometric constraints is one
@@
 The experiments we have conducted are also likely to have too simplistic
-semantics to let complex reasoning needs to emerge. Scenarios with more complex
+semantics to let complex reasoning needs emerge. Scenarios with more complex
 semantics would be desirable to better stress the expressiveness and inference
@@
 application of reasoning in our architecture, where reasoning related
-to decision making mostly happen outside the knowledge representation system.
+to decision making mostly happens outside the knowledge representation system.
 
@@
 This transformation, both from the technical point of view (the {\sc Oro}
-server, the ontologies, the bindings, etc.), and as a more subtle change in the
+server, the ontologies, the bindings, etc.) and as a more subtle change in the
 practises related to the development of robotic components, is probably the
@@
 
-Communication between the robot's module is now conceived in term of what are
+Communication between the robot's modules is now conceived in terms of what are
 the \emph{semantics} of the information flows, instead of a simple
 compatibility of interfaces. When defining the frontiers of a robotic
-component, we do not think anymore only in term of \emph{is the interface complete
-and self-contained}, but also in term of \emph{is the semantic complete and
-consistent?}. This allow a deeper, more correct modularity: two modules that
+component, we do not think anymore only in terms of \emph{is the interface complete
+and self-contained}, but also in terms of \emph{is the semantic complete and
+consistent?}. This allows a deeper, more correct modularity: two modules that
 share the same, well-defined semantic can be confidently exchanged. When we
@@
 
-Even more, this architecture also contribute to bring closer robotics and
+Even more, this architecture also contributes to bring closer robotics and
 cognitive psychology: it provides clear entry points to implement some
@@
 
-One of the difficult challenge of cognitive robotics, however, is to overcome
+One of the difficult challenges of cognitive robotics, however, is to overcome
 the relative isolation of each of these communities to build an autonomous
